---
title: "Climate Anomalies Around Hot Days: Post-Defense Analysis"
author: "Penny Hunter"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  warning = FALSE,
  message = FALSE
)
source("~/Desktop/ASCI-GitHub/Masters-Thesis/setup.R")
```

---

## Overview

This script constructs daily climate anomalies relative to a smoothed
day-of-year climatology and estimates how soil moisture, temperature, VPD,
and precipitation in the 30-day window around hot days (Tmax ≥ 30°C).
Inference is based on a state-year block bootstrap (1,000 replications) to
account for spatial and temporal dependence.

Pipeline:

1. Load LDAS soil moisture and PRISM climate data
2. Merge all variables into a single daily county-level panel
3. Compute smoothed DOY climatology and daily anomalies
4. Identify hot days and run the bootstrap
5. Plot results

---

## 1. Load LDAS Soil Moisture Data

LDAS soil moisture is available at three depth layers. Files are filtered to
the historical period (1980–2018).

```{r load-ldas}
load_ldas_data <- function(folder_path) {
  all_files <- list.files(path = folder_path, pattern = "\\.csv$", full.names = TRUE)
  files_to_load <- grep(
    pattern = "_19[8-9][0-9]\\.csv$|_200[0-9]\\.csv$|_201[0-8]\\.csv$",
    x       = all_files,
    value   = TRUE
  )
  lapply(files_to_load, read.csv) |> dplyr::bind_rows()
}

ldas_sm_0_10   <- load_ldas_data("input_data/LDAS_SM_0_10")
ldas_sm_10_40  <- load_ldas_data("input_data/LDAS_SM_10_40")
ldas_sm_40_100 <- load_ldas_data("input_data/LDAS_SM_40_100")
```

---

## 2. Load PRISM Climate Data

```{r load-prism}
ppt_county_full     <- readRDS("input_data/prism_data_full/ppt_county_full.rds")
tmin_county_full    <- readRDS("input_data/prism_data_full/tmin_county_full.rds")
tmax_county_full    <- readRDS("input_data/prism_data_full/tmax_county_full.rds")
vpdmin_county_full  <- readRDS("input_data/prism_data_full/vpdmin_county_full.rds")
vpdmean_county_full <- readRDS("input_data/prism_data_full/vpdmean_county_full.rds")
vpdmax_county_full  <- readRDS("input_data/prism_data_full/vpdmax_county_full.rds")
```

---

## 3. Prepare and Merge All Variables

Drop redundant identifier columns from PRISM datasets, standardize LDAS
column names and GEOID formatting, then join all variables into a single
daily panel.

```{r prep-and-merge}
# Drop redundant columns from PRISM datasets
prism_datasets <- list(
  tmin_county_full, ppt_county_full,
  vpdmax_county_full, vpdmin_county_full, vpdmean_county_full
)
dataset_names  <- c(
  "tmin_county_full", "ppt_county_full",
  "vpdmax_county_full", "vpdmin_county_full", "vpdmean_county_full"
)
cols_to_remove <- c("county_name", "county_count", "Year", "Month", "Day")

for (i in seq_along(prism_datasets)) {
  df <- prism_datasets[[i]]
  df <- df[, !(names(df) %in% cols_to_remove)]
  assign(dataset_names[i], df, envir = .GlobalEnv)
}

# Standardize LDAS layers
prep_ldas <- function(df, col_name) {
  df %>%
    mutate(
      time  = as.Date(time),
      GEOID = stringr::str_pad(as.character(GEOID), width = 5, pad = "0"),
      !!col_name := var
    ) %>%
    select(GEOID, time, all_of(col_name))
}

ldas_sm_0_10   <- prep_ldas(ldas_sm_0_10,   "SM_0_10")
ldas_sm_10_40  <- prep_ldas(ldas_sm_10_40,  "SM_10_40")
ldas_sm_40_100 <- prep_ldas(ldas_sm_40_100, "SM_40_100")

# Merge all variables
merged_prism_full <- tmax_county_full %>%
  full_join(tmin_county_full,    by = c("time", "GEOID")) %>%
  full_join(ppt_county_full,     by = c("time", "GEOID")) %>%
  full_join(vpdmax_county_full,  by = c("time", "GEOID")) %>%
  full_join(vpdmean_county_full, by = c("time", "GEOID")) %>%
  full_join(vpdmin_county_full,  by = c("time", "GEOID")) %>%
  full_join(ldas_sm_0_10,        by = c("time", "GEOID")) %>%
  full_join(ldas_sm_10_40,       by = c("time", "GEOID")) %>%
  full_join(ldas_sm_40_100,      by = c("time", "GEOID"))

merged_prism_full_1 <- merged_prism_full

rm(merged_prism_full, ppt_county_full, vpdmax_county_full, vpdmean_county_full,
   vpdmin_county_full, tmin_county_full, tmax_county_full,
   ldas_sm_0_10, ldas_sm_10_40, ldas_sm_40_100, prism_datasets, df)

saveRDS(
  merged_prism_full_1,
  "/Users/pennyhunter/Desktop/ASCI-GitHub/Masters-Thesis/input_data/merged_prism_full_1_wldas.RDS"
)
# merged_prism_full_1 <- readRDS("input_data/merged_prism_full_1_wldas.RDS")
```

```{r filter-counties}
# Restrict to counties present in the yield analysis
counties_combined   <- readRDS("input_data/counties_combined.rds")
merged_prism_full_1 <- merged_prism_full_1 %>%
  filter(GEOID %in% counties_combined$fips)
rm(counties_combined)
```

---

## 4. Compute DOY Climatology and Daily Anomalies

A smoothed day-of-year (DOY) climatology is computed for each county using a
±7-day centred running mean. Anomalies are then calculated as observed minus
climatological values.

```{r compute-anomalies}
library(slider)

clim_vars <- c(
  "prism_vpdmax", "prism_vpdmin", "prism_vpdmean",
  "prism_ppt", "prism_tmax", "prism_tmin",
  "SM_0_10", "SM_10_40", "SM_40_100"
)

merged_prism_full_1 <- merged_prism_full_1 %>%
  mutate(doy = lubridate::yday(time))

# Raw DOY climatology (county × DOY mean across all years)
doy_clim <- merged_prism_full_1 %>%
  group_by(GEOID, doy) %>%
  summarise(across(all_of(clim_vars),
                   ~ mean(.x, na.rm = TRUE),
                   .names = "clim_{.col}"),
            .groups = "drop")

# Smooth with a ±7-day centred running mean
doy_clim <- doy_clim %>%
  group_by(GEOID) %>%
  arrange(doy, .by_group = TRUE) %>%
  mutate(across(
    starts_with("clim_"),
    ~ slide_dbl(.x, ~ mean(.x, na.rm = TRUE),
                .before = 7, .after = 7, .complete = FALSE)
  )) %>%
  ungroup()

# Compute anomalies
merged_prism_full_1 <- merged_prism_full_1 %>%
  left_join(doy_clim, by = c("GEOID", "doy")) %>%
  mutate(
    anomaly_vpdmax    = prism_vpdmax  - clim_prism_vpdmax,
    anomaly_vpdmin    = prism_vpdmin  - clim_prism_vpdmin,
    anomaly_vpdmean   = prism_vpdmean - clim_prism_vpdmean,
    anomaly_ppt       = prism_ppt     - clim_prism_ppt,
    anomaly_tmax      = prism_tmax    - clim_prism_tmax,
    anomaly_tmin      = prism_tmin    - clim_prism_tmin,
    anomaly_SM_0_10   = SM_0_10       - clim_SM_0_10,
    anomaly_SM_10_40  = SM_10_40      - clim_SM_10_40,
    anomaly_SM_40_100 = SM_40_100     - clim_SM_40_100
  )

rm(doy_clim)
```

---

## 5. Construct Daily GDD, KDD, and Hot Day Indicator

```{r gdd-kdd}
kdd_threshold <- 30
gdd_base      <- 10
gdd_cap       <- 30

Tf <- 0.44 - 0.46 * sin((pi * seq(0.5, 23, 1)) / 12 + 0.9) +
      0.11 * sin((2 * pi * seq(0.5, 23, 1)) / 12 + 0.9)

calculate_kdd <- function(T_max, T_min, thr, Tf) {
  hrlyT <- T_max * Tf + T_min * (1 - Tf)
  sum(pmax(hrlyT - thr, 0)) / 24
}

calculate_gdd <- function(T_max, T_min, Tf, base = gdd_base, cap = gdd_cap) {
  if (T_min < base || T_max > cap) return(0)
  hrlyT <- T_max * Tf + T_min * (1 - Tf)
  sum(pmax(hrlyT - base, 0)) / 24
}

merged_prism_full_1$KDD <- mapply(
  calculate_kdd,
  T_max    = merged_prism_full_1$prism_tmax,
  T_min    = merged_prism_full_1$prism_tmin,
  MoreArgs = list(thr = kdd_threshold, Tf = Tf)
)

merged_prism_full_1$GDD <- mapply(
  calculate_gdd,
  T_max    = merged_prism_full_1$prism_tmax,
  T_min    = merged_prism_full_1$prism_tmin,
  MoreArgs = list(Tf = Tf)
)

merged_prism_full_1 <- merged_prism_full_1 %>%
  mutate(
    year      = lubridate::year(time),
    binary_30 = as.integer(!is.na(prism_tmax) & prism_tmax >= 30),
    GEOID     = stringr::str_pad(as.character(GEOID), width = 5, pad = "0"),
    state     = substr(GEOID, 1, 2)
  ) %>%
  group_by(GEOID, year) %>%
  mutate(
    KDD_sum_yearly = sum(KDD, na.rm = TRUE),
    GDD_sum_yearly = sum(GDD, na.rm = TRUE)
  ) %>%
  ungroup()

saveRDS(
  merged_prism_full_1,
  "/Users/pennyhunter/Desktop/ASCI-GitHub/Masters-Thesis/input_data/merged_prism_full_1_wldas.RDS"
)
# merged_prism_full_1 <- readRDS("input_data/merged_prism_full_1_wldas.RDS")
```

---

## 6. State-Year Block Bootstrap

For each anomaly variable, KDD-weighted mean anomalies are computed in a
±30-day window around each hot day. Confidence intervals are derived from
1,000 state-year block bootstrap replications.

```{r bootstrap, cache=TRUE}
library(data.table)

setDT(merged_prism_full_1)
merged_all <- merged_prism_full_1

hot_days <- merged_all[binary_30 == 1, .(GEOID, state, day0_date = time, KDD)]
hot_days[, year := year(day0_date)]

anomaly_vars <- c(
  "anomaly_tmax",
  "anomaly_vpdmax",
  "anomaly_vpdmean",
  "anomaly_vpdmin",
  "anomaly_ppt",
  "anomaly_SM_0_10",
  "anomaly_SM_10_40",
  "anomaly_SM_40_100"
)

day_offsets        <- -30:30
n_boot             <- 1000
set.seed(123)

unique_state_year <- unique(hot_days[, .(state, year)])
all_results       <- list()

for (var in anomaly_vars) {
  cat("\nProcessing:", var, "\n")

  anomaly_lookup <- merged_all[, .(GEOID, time, anomaly_value = get(var))]

  bootstrap_mat <- matrix(NA_real_, nrow = n_boot, ncol = length(day_offsets))
  colnames(bootstrap_mat) <- paste0("d", day_offsets)

  for (i in seq_len(n_boot)) {
    if (i %% 50 == 0) cat("  Bootstrap", i, "of", n_boot, "\n")

    sampled_blocks <- unique_state_year[sample(.N, replace = TRUE)]

    hot_sample <- rbindlist(lapply(seq_len(nrow(sampled_blocks)), function(j) {
      hot_days[state == sampled_blocks$state[j] & year == sampled_blocks$year[j]]
    }))

    boot_windows <- hot_sample[, .(offset = day_offsets), by = .(GEOID, day0_date, KDD)]
    boot_windows[, window_date := day0_date + offset]

    boot_window_data <- merge(
      boot_windows, anomaly_lookup,
      by.x = c("GEOID", "window_date"), by.y = c("GEOID", "time"),
      all.x = TRUE
    )

    boot_window_data[, weighted_anomaly := anomaly_value * KDD]
    total_kdd <- sum(hot_sample$KDD, na.rm = TRUE)

    avg_by_offset <- boot_window_data[
      !is.na(weighted_anomaly),
      .(avg = sum(weighted_anomaly, na.rm = TRUE) / total_kdd),
      by = offset
    ]
    setkey(avg_by_offset, offset)

    bootstrap_mat[i, match(avg_by_offset$offset, day_offsets)] <- avg_by_offset$avg
  }

  all_results[[var]] <- data.table(
    variable = var,
    day      = day_offsets,
    mean     = colMeans(bootstrap_mat, na.rm = TRUE),
    lower    = apply(bootstrap_mat, 2, quantile, probs = 0.025, na.rm = TRUE),
    upper    = apply(bootstrap_mat, 2, quantile, probs = 0.975, na.rm = TRUE)
  )
}

all_boot <- rbindlist(all_results)

saveRDS(
  all_boot,
  "/Users/pennyhunter/Desktop/ASCI-GitHub/Masters-Thesis/output_data/all_boot_STATE_YEAR.RDS"
)
# all_boot <- readRDS("output_data/all_boot_STATE_YEAR.RDS")
```

---

## 7. Plot Bootstrap Results

Individual and combined panel figures are saved at 600 dpi. Soil moisture
panels share a common y-axis. Panels are labelled a–f.

```{r plot-setup}
library(patchwork)

# Variable metadata: y-axis labels
var_meta <- list(
  anomaly_ppt       = list(label = "Precipitation Anomaly (mm)"),
  anomaly_SM_0_10   = list(label = "Soil Moisture 0–10 cm Anomaly (m³/m³)"),
  anomaly_SM_10_40  = list(label = "Soil Moisture 10–40 cm Anomaly (m³/m³)"),
  anomaly_SM_40_100 = list(label = "Soil Moisture 40–100 cm Anomaly (m³/m³)"),
  anomaly_tmax      = list(label = "Max Temperature Anomaly (°C)"),
  anomaly_vpdmax    = list(label = "VPD Maximum Anomaly (hPa)")
)

# Plot order and panel letters
plot_order   <- c(
  "anomaly_ppt",      "anomaly_SM_0_10",
  "anomaly_SM_10_40", "anomaly_SM_40_100",
  "anomaly_tmax",     "anomaly_vpdmax"
)
plot_letters <- setNames(letters[seq_along(plot_order)], plot_order)

# Shared y-axis limits for soil moisture panels
sm_vars     <- c("anomaly_SM_0_10", "anomaly_SM_10_40", "anomaly_SM_40_100")
all_boot_dt <- as.data.table(all_boot)
sm_data     <- all_boot_dt[variable %in% sm_vars]
sm_ylim     <- c(min(sm_data$lower, na.rm = TRUE), max(sm_data$upper, na.rm = TRUE))

bottom_row_vars <- c("anomaly_tmax", "anomaly_vpdmax")
```

```{r plot-function}
make_plot <- function(var, letter, ylim_fixed = NULL, show_x_axis = FALSE) {
  dat    <- all_boot_dt[variable == var]
  ylabel <- var_meta[[var]]$label

  p <- ggplot(dat, aes(x = day)) +
    geom_vline(xintercept = 0, linetype = "dotted",
               color = "gray50", linewidth = 0.6) +
    geom_ribbon(aes(ymin = lower, ymax = upper),
                fill = "#8FBC8F", alpha = 0.45) +
    geom_line(aes(y = mean), color = "darkgreen", linewidth = 1.1) +
    geom_hline(yintercept = 0, linetype = "dashed",
               color = "gray50", linewidth = 0.5) +
    labs(
      title = letter,
      x     = if (show_x_axis) expression(bold("Days from Hot Day")) else NULL,
      y     = ylabel
    ) +
    theme(
      panel.grid        = element_blank(),
      panel.border      = element_rect(color = "black", fill = NA, linewidth = 0.5),
      plot.title        = element_text(size = 11, face = "bold", hjust = 0),
      axis.title.y      = element_text(size = 10, face = "bold"),
      axis.text         = element_text(size = 9),
      axis.title.x      = if (!show_x_axis) element_blank()
                          else element_text(size = 11, face = "bold"),
      axis.ticks        = element_line(color = "black", linewidth = 0.4),
      axis.ticks.length = unit(3, "pt")
    ) +
    scale_x_continuous(expand = c(0, 0))

  if (!is.null(ylim_fixed)) p <- p + scale_y_continuous(limits = ylim_fixed)
  return(p)
}
```

```{r save-individual-plots}
for (var in plot_order) {
  p <- make_plot(
    var,
    letter      = plot_letters[var],
    ylim_fixed  = if (var %in% sm_vars) sm_ylim else NULL,
    show_x_axis = TRUE
  )
  fname <- paste0("output_data/", var, "_bootstrap_STATE_YEAR.png")
  ggsave(fname, p, width = 7.5, height = 4, dpi = 600, units = "in")
  cat("Saved:", fname, "\n")
}
```

```{r combined-panel, fig.width=7.5, fig.height=10}
plot_list <- lapply(plot_order, function(var) {
  make_plot(
    var,
    letter      = plot_letters[var],
    ylim_fixed  = if (var %in% sm_vars) sm_ylim else NULL,
    show_x_axis = var %in% bottom_row_vars
  )
})

p_all <- wrap_plots(plot_list, ncol = 2)
print(p_all)

ggsave(
  "output_data/ALL_VARIABLES_bootstrap_STATE_YEAR.png",
  p_all, width = 7.5, height = 10, dpi = 600, units = "in"
)
cat("Saved: ALL_VARIABLES_bootstrap_STATE_YEAR.png\n")
```
