---
title: "Bootstrap Yield Projections Under Climate Change"
author: "Penny Hunter"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(lubridate)
library(ggplot2)
library(arrow)
```

# Overview
# For each of the 4 models (corn irrigated, corn rainfed, soy irrigated, soy rainfed):
# 1. Run 1000 bootstrap iterations using state-year block resampling on historical PRISM data
# 2. For each bootstrap iteration x 5 GCMs = 5000 yield trajectories per crop model
# 3. Each trajectory = one bootstrap model fit on resampled historical data,
#    predicted through one GCM with year frozen at 2018
# 4. 2.5% and 97.5% quantiles across 5000 trajectories = 95% CI band
# 5. Mean across 5000 trajectories = central estimate line

# ============================================================
# 1. LOAD DATA AND MODELS
# ============================================================

```{r load-data}
# Load historical growing season data (used for bootstrap resampling)
maize_gs     <- readRDS("input_data/maize_gs")
soybeans_gs  <- readRDS("input_data/soybeans_gs")

# Ensure county_factor uses fips (consistent with model training)
maize_gs <- maize_gs %>%
  mutate(county_factor = factor(fips))

soybeans_gs <- soybeans_gs %>%
  mutate(county_factor = factor(fips))

# Load trained models (used as reference for factor levels and formula)
model_maize_irrig <- readRDS("input_data/model_maize_irrig.rds")
model_maize_rain  <- readRDS("input_data/model_maize_rain.rds")
model_soy_irrig   <- readRDS("input_data/model_soy_irrig.rds")
model_soy_rain    <- readRDS("input_data/model_soy_rain.rds")

# Load county lists for CMIP cropping
maize_counties   <- as.character(unique(maize_gs$fips))
soybean_counties <- as.character(unique(soybeans_gs$fips))
```

# ============================================================
# 2. CMIP DATA LOADING
# ============================================================

```{r load-cmip}
# Function to crop CMIP parquet to training counties
crop_cmip_to_training <- function(parquet_path, counties) {
  ds <- open_dataset(parquet_path)
  vars_needed <- c("GEOID", "time", "tasmax", "tasmin",
                   "pr", "daily_mean_vpd_tdew_bc_temp")
  ds_small <- ds |>
    select(any_of(vars_needed)) |>
    filter(GEOID %in% counties)
  collect(ds_small)
}

# CMIP file paths (SSP2-4.5)
cmip_files <- list(
  GFDL  = "input_data/all_GFDL-ESM4_ssp245_bc.parquet",
  IPSL  = "input_data/all_IPSL-CM6A-LR_ssp245_bc.parquet",
  MPI   = "input_data/all_MPI-ESM1-2-HR_ssp245_bc.parquet",
  MRI   = "input_data/all_MRI-ESM2-0_ssp245_bc.parquet",
  UKESM = "input_data/all_UKESM1-0-LL_ssp245_bc.parquet"
)

# Load and crop all GCMs
cmip_cropped <- list()

for (m in names(cmip_files)) {
  message("Cropping CMIP data for GCM: ", m)
  cmip_cropped[[m]] <- list(
    maize = crop_cmip_to_training(cmip_files[[m]], maize_counties),
    soy   = crop_cmip_to_training(cmip_files[[m]], soybean_counties)
  )
  # Pad GEOID to 5 digits
  cmip_cropped[[m]]$maize <- cmip_cropped[[m]]$maize %>%
    mutate(GEOID = sprintf("%05s", GEOID))
  cmip_cropped[[m]]$soy <- cmip_cropped[[m]]$soy %>%
    mutate(GEOID = sprintf("%05s", GEOID))
}
```

# ============================================================
# 3. KDD / GDD FUNCTIONS (unchanged from future projection script)
# ============================================================

```{r kdd-gdd-functions}
kdd_threshold <- 30
gdd_base      <- 10
gdd_cap       <- 30

# Hourly temperature profile
Tf <- 0.44 - 0.46 * sin((pi * seq(0.5, 23, 1)) / 12 + 0.9) +
      0.11 * sin((2 * pi * seq(0.5, 23, 1)) / 12 + 0.9)

# Vectorised daily KDD and GDD using matrix multiplication
compute_kdd_gdd <- function(df) {
  hrlyT <- outer(df$tasmax, Tf) * 1 + outer(df$tasmin, (1 - Tf))

  KDD_mat   <- pmax(hrlyT - kdd_threshold, 0)
  KDD_daily <- rowMeans(KDD_mat)

  hrlyT_GDD <- pmin(pmax(hrlyT, gdd_base), gdd_cap)
  GDD_mat   <- pmax(hrlyT_GDD - gdd_base, 0)
  GDD_daily <- rowMeans(GDD_mat)

  df$GDD_daily <- GDD_daily
  df$KDD_daily <- KDD_daily
  df
}

# Aggregate daily CMIP data to county-year
aggregate_county_year <- function(df) {
  df %>%
    mutate(year = year(time)) %>%
    group_by(GEOID, year) %>%
    summarise(
      GDD_sum_yearly = sum(GDD_daily, na.rm = TRUE),
      KDD_sum_yearly = sum(KDD_daily, na.rm = TRUE),
      prism_ppt      = sum(pr,        na.rm = TRUE),
      vpd_mean       = mean(daily_mean_vpd_tdew_bc_temp, na.rm = TRUE),
      .groups = "drop"
    )
}
```

# ============================================================
# 4. PRE-PROCESS ALL CMIP DATA ONCE
# (compute KDD/GDD and aggregate before the bootstrap loop
#  so we are not repeating expensive operations 1000x per GCM)
# ============================================================

```{r preprocess-cmip}
frozen_year <- 2018

# Process one GCM x crop combination into a prediction-ready data frame
prepare_cmip_for_prediction <- function(cmip_df, county_levels) {
  cmip_df %>%
    compute_kdd_gdd() %>%
    aggregate_county_year() %>%
    mutate(
      GEOID         = sprintf("%05s", GEOID),
      cip_year      = year,       # keep actual CMIP year for plotting
      year          = frozen_year # freeze year to remove technology trend
    ) %>%
    filter(GEOID %in% county_levels) %>%
    mutate(county_factor = factor(GEOID, levels = county_levels))
}

# Extract county levels from each trained model
maize_county_levels <- levels(model_maize_irrig$model$county_factor)
soy_county_levels   <- levels(model_soy_irrig$model$county_factor)

# Pre-process all 5 GCMs for maize and soy
# Result: cmip_prepped[[gcm]]$maize  and  cmip_prepped[[gcm]]$soy
cmip_prepped <- list()

for (m in names(cmip_cropped)) {
  message("Pre-processing CMIP: ", m)
  cmip_prepped[[m]] <- list(
    maize = prepare_cmip_for_prediction(cmip_cropped[[m]]$maize, maize_county_levels),
    soy   = prepare_cmip_for_prediction(cmip_cropped[[m]]$soy,   soy_county_levels)
  )
}

# Free raw CMIP data from memory
rm(cmip_cropped)
gc()
```

# ============================================================
# 5. STATE-YEAR BLOCK BOOTSTRAP FUNCTION
# (unchanged logic from historical script)
# ============================================================

```{r bootstrap-function}
# Bootstrap a single model formula on historical data using state-year blocks.
# Returns a fitted lm object from one resampled dataset.
bootstrap_one_model <- function(data, formula,
                                state_var = "state_ansi",
                                year_var  = "year") {

  data <- data %>%
    mutate(state_year = interaction(.data[[state_var]],
                                    .data[[year_var]], drop = TRUE))

  unique_blocks  <- unique(data$state_year)
  sampled_blocks <- sample(unique_blocks,
                           size    = length(unique_blocks),
                           replace = TRUE)

  boot_sample <- do.call(rbind, lapply(sampled_blocks, function(b) {
    data[data$state_year == b, ]
  }))

  tryCatch(
    lm(formula, data = boot_sample),
    error = function(e) NULL   # return NULL if model fails to fit
  )
}
```

# ============================================================
# 6. MAIN BOOTSTRAP LOOP
# For each crop model:
#   - 1000 bootstrap iterations
#   - each iteration predicts through all 5 GCMs
#   - store mean county yield per year for each bootstrap x GCM
#   - total trajectories per crop model = 1000 x 5 = 5000
# ============================================================

```{r bootstrap-loop}
set.seed(123)
n_boot <- 1000

# Define the four model configurations
model_configs <- list(
  corn_irrig = list(
    data      = maize_gs,
    cmip_crop = "maize",
    formula   = as.formula(
      "irr ~ KDD_sum_yearly + GDD_sum_yearly + prism_ppt +
       I(prism_ppt^2) + county_factor + factor(year)"
    ),
    label     = "Maize Irrigated"
  ),
  corn_rain = list(
    data      = maize_gs,
    cmip_crop = "maize",
    formula   = as.formula(
      "rainfed ~ KDD_sum_yearly + GDD_sum_yearly + prism_ppt +
       I(prism_ppt^2) + county_factor + factor(year)"
    ),
    label     = "Maize Rainfed"
  ),
  soy_irrig = list(
    data      = soybeans_gs,
    cmip_crop = "soy",
    formula   = as.formula(
      "irr ~ KDD_sum_yearly + GDD_sum_yearly + prism_ppt +
       I(prism_ppt^2) + county_factor + factor(year)"
    ),
    label     = "Soy Irrigated"
  ),
  soy_rain = list(
    data      = soybeans_gs,
    cmip_crop = "soy",
    formula   = as.formula(
      "rainfed ~ KDD_sum_yearly + GDD_sum_yearly + prism_ppt +
       I(prism_ppt^2) + county_factor + factor(year)"
    ),
    label     = "Soy Rainfed"
  )
)

# Storage: for each model config, a list of 5000 year-mean trajectories
# Each element is a data frame with columns: cip_year, mean_yield
all_trajectories <- list()

for (config_name in names(model_configs)) {
  cfg <- model_configs[[config_name]]
  message("\n===== Starting bootstrap for: ", cfg$label, " =====")

  trajectories <- vector("list", n_boot * length(cmip_prepped))
  traj_index   <- 1

  for (b in seq_len(n_boot)) {

    if (b %% 100 == 0) message("  Bootstrap iteration: ", b, " / ", n_boot)

    # Fit one bootstrap model on resampled historical data
    boot_model <- bootstrap_one_model(cfg$data, cfg$formula)

    # Skip this iteration if model failed to fit
    if (is.null(boot_model)) next

    # Predict through all 5 GCMs using this bootstrap model
    for (gcm_name in names(cmip_prepped)) {

      cmip_df <- cmip_prepped[[gcm_name]][[cfg$cmip_crop]]

      # Align county_factor levels to the bootstrap model
      # NOTE: bootstrap model may have seen different county factor levels
      #       so we filter to counties present in both
      valid_counties <- intersect(
        as.character(cmip_df$GEOID),
        names(coef(boot_model))[grepl("county_factor", names(coef(boot_model)))] %>%
          gsub("county_factor", "", .)
      )

      # Safer approach: align factor levels to boot_model levels
      boot_county_levels <- tryCatch(
        levels(boot_model$model$county_factor),
        error = function(e) NULL
      )

      if (is.null(boot_county_levels)) next

      pred_df <- cmip_df %>%
        filter(GEOID %in% boot_county_levels) %>%
        mutate(county_factor = factor(GEOID, levels = boot_county_levels))

      if (nrow(pred_df) == 0) next

      # Predict yields, cap negatives at 0
      preds <- tryCatch(
        predict(boot_model, newdata = pred_df),
        error = function(e) NULL
      )

      if (is.null(preds)) next

      pred_df$predicted_yield <- pmax(preds, 0)

      # Summarise to mean yield per year across counties
      year_means <- pred_df %>%
        group_by(cip_year) %>%
        summarise(mean_yield = mean(predicted_yield, na.rm = TRUE),
                  .groups = "drop") %>%
        filter(cip_year != 2014) %>%   # drop spin-up year
        mutate(boot_iter = b, gcm = gcm_name)

      trajectories[[traj_index]] <- year_means
      traj_index <- traj_index + 1
    }
  }

  # Combine all trajectories for this config (up to 5000 rows per year)
  all_trajectories[[config_name]] <- bind_rows(trajectories)
  message("  Total trajectories stored: ",
          n_distinct(
            paste(all_trajectories[[config_name]]$boot_iter,
                  all_trajectories[[config_name]]$gcm)
          ))
}

# Save trajectories for re-use without re-running
saveRDS(all_trajectories, "input_data/bootstrap_trajectories.rds")
message("Bootstrap complete. Trajectories saved.")
```

# ============================================================
# 7. SUMMARISE TRAJECTORIES INTO CI BANDS
# ============================================================
```{r summarise-trajectories}

#all_trajectories <- readRDS("input_data/bootstrap_trajectories.rds")
# CO2 fertilization setup (Hultgren et al. 2025 / Moore et al. 2017)
co2_ssp245 <- data.frame(
  year    = 2015:2100,
  co2_ppm = approx(
    x    = c(2015, 2030, 2050, 2070, 2100),
    y    = c(400,  423,  457,  487,  538),
    xout = 2015:2100
  )$y
)

co2_fertilization <- function(co2_ppm, beta, A, baseline_ppm = 400) {
  delta <- co2_ppm - baseline_ppm
  (beta * delta / (delta + A)) / 100
}

co2_ssp245 <- co2_ssp245 %>%
  mutate(
    co2_mult_maize = exp(co2_fertilization(co2_ppm, beta = 10.82, A = 50)),
    co2_mult_soy   = exp(co2_fertilization(co2_ppm, beta = 17.20, A = 100))
  )

# Summarise trajectories and apply CO2 fertilization (yields already in MT/ha)
summarise_trajectories <- function(traj_df) {
  traj_df %>%
    rename(traj_yield = mean_yield) %>%
    group_by(cip_year) %>%
    summarise(
      mean_yield = mean(traj_yield,            na.rm = TRUE),
      lower_95   = quantile(traj_yield, 0.025, na.rm = TRUE),
      upper_95   = quantile(traj_yield, 0.975, na.rm = TRUE),
      n_traj     = n(),
      .groups    = "drop"
    ) %>%
    rename(year = cip_year)
}

corn_irrig_summary <- summarise_trajectories(all_trajectories$corn_irrig) %>%
  mutate(Type = "Irrigated") %>%
  left_join(co2_ssp245 %>% select(year, co2_mult_maize), by = "year") %>%
  mutate(across(c(mean_yield, lower_95, upper_95), ~ . * co2_mult_maize)) %>%
  select(-co2_mult_maize)

corn_rain_summary <- summarise_trajectories(all_trajectories$corn_rain) %>%
  mutate(Type = "Rainfed") %>%
  left_join(co2_ssp245 %>% select(year, co2_mult_maize), by = "year") %>%
  mutate(across(c(mean_yield, lower_95, upper_95), ~ . * co2_mult_maize)) %>%
  select(-co2_mult_maize)

soy_irrig_summary <- summarise_trajectories(all_trajectories$soy_irrig) %>%
  mutate(Type = "Irrigated") %>%
  left_join(co2_ssp245 %>% select(year, co2_mult_soy), by = "year") %>%
  mutate(across(c(mean_yield, lower_95, upper_95), ~ . * co2_mult_soy)) %>%
  select(-co2_mult_soy)

soy_rain_summary <- summarise_trajectories(all_trajectories$soy_rain) %>%
  mutate(Type = "Rainfed") %>%
  left_join(co2_ssp245 %>% select(year, co2_mult_soy), by = "year") %>%
  mutate(across(c(mean_yield, lower_95, upper_95), ~ . * co2_mult_soy)) %>%
  select(-co2_mult_soy)

corn_summary <- bind_rows(corn_irrig_summary, corn_rain_summary)
soy_summary  <- bind_rows(soy_irrig_summary,  soy_rain_summary)
```

# ============================================================
# 8. PLOT 1: TIME SERIES WITH 95% CI BOOTSTRAP RIBBON
# ============================================================

```{r plot-ci}
cols <- c("Irrigated" = "#1f78b4", "Rainfed" = "#33a02c")

make_plot_bootstrap_ci <- function(df, crop_name, file_name) {

  p <- ggplot(df, aes(x = year, y = mean_yield,
                      color = Type, fill = Type)) +
    geom_ribbon(aes(ymin = lower_95, ymax = upper_95, group = Type),
                alpha = 0.2, color = NA) +
    geom_line(aes(group = Type), size = 1.2) +
    scale_color_manual(
      values = cols,
      labels = c("Irrigated" = "IRRIGATED", "Rainfed" = "RAINFED")
    ) +
    scale_fill_manual(
      values = cols,
      labels = c("Irrigated" = "IRRIGATED", "Rainfed" = "RAINFED")
    ) +
    scale_x_continuous(
      expand = c(0, 0),
      breaks = c(min(df$year), 2050, 2100)
    ) +
    scale_y_continuous(expand = c(0, 0)) +
    theme_classic(base_size = 16) +
    theme(
      legend.position  = "bottom",
      legend.title     = element_blank(),
      legend.text      = element_text(size = 14),
      axis.title       = element_text(face = "bold", size = 16),
      axis.text        = element_text(size = 14),
      plot.title       = element_text(face = "bold", size = 18, hjust = 0.5),
      axis.line        = element_line(size = 0.8),
      panel.border     = element_rect(color = "black", fill = NA, size = 0.8)
    ) +
    labs(
      x     = "Year",
      y     = "Predicted Yield (MT/ha)",
      title = paste0(crop_name, " — Bootstrap 95% CI (n = 5,000 trajectories)")
    )

  ggsave(file_name, p, width = 8, height = 4.5, dpi = 300)
  p
}

corn_ci_plot <- make_plot_bootstrap_ci(
  corn_summary, "Maize", "output_data/corn_bootstrap_ci.png"
)

soy_ci_plot  <- make_plot_bootstrap_ci(
  soy_summary,  "Soy",  "output_data/soy_bootstrap_ci.png"
)

corn_ci_plot
soy_ci_plot
```



# ============================================================
# 9. YIELD CHANGE TABLE (first 20 vs last 20 years of projections)
# ============================================================

```{r yield-change-table}
compute_yield_change_bootstrap <- function(summary_df, crop, type_label) {
  df <- summary_df %>% filter(Type == type_label) %>% arrange(year)

  years      <- sort(unique(df$year))
  first_years <- years[1:20]
  last_years  <- tail(years, 20)

  first_mean <- df %>%
    filter(year %in% first_years) %>%
    summarise(m = mean(mean_yield, na.rm = TRUE)) %>%
    pull(m)

  last_mean  <- df %>%
    filter(year %in% last_years) %>%
    summarise(m = mean(mean_yield, na.rm = TRUE)) %>%
    pull(m)

  data.frame(
    Crop        = crop,
    Type        = type_label,
    First20_mean = round(first_mean, 2),
    Last20_mean  = round(last_mean,  2),
    Abs_change   = round(last_mean - first_mean, 2),
    Pct_change   = round(((last_mean - first_mean) / first_mean) * 100, 2)
  )
}

yield_change_table <- bind_rows(
  compute_yield_change_bootstrap(corn_summary, "Maize", "Irrigated"),
  compute_yield_change_bootstrap(corn_summary, "Maize", "Rainfed"),
  compute_yield_change_bootstrap(soy_summary,  "Soy",  "Irrigated"),
  compute_yield_change_bootstrap(soy_summary,  "Soy",  "Rainfed")
)

print(yield_change_table)
```

# ============================================================
# 10. SPAGHETTI PLOT: All 5000 trajectories + central mean line
# ============================================================

```{r yield-spaghetti-plot}

make_spaghetti_plot <- function(traj_df, summary_df, crop_name, file_name) {
  
  cols <- c("Irrigated" = "#1f78b4", "Rainfed" = "#33a02c")
  
  # Add Type label to raw trajectories using config name
  # traj_df must already have a Type column — bind irrigated and rainfed first
  
  p <- ggplot() +
    # All 5000 individual trajectory lines — very transparent
    geom_line(
      data = traj_df,
      aes(x = cip_year, y = mean_yield, 
          group = interaction(boot_iter, gcm, Type),
          color = Type),
      alpha = 0.005,
      linewidth = 0.3
    ) +
    # Central mean lines on top — thicker, same colors
    geom_line(
      data = summary_df,
      aes(x = year, y = mean_yield,
          color = Type, group = Type),
      linewidth = 1.8
    ) +
    scale_color_manual(
      values = cols,
      labels = c("Irrigated" = "IRRIGATED", "Rainfed" = "RAINFED")
    ) +
    scale_x_continuous(
      expand = c(0, 0),
      breaks = c(2015, 2050, 2100)
    ) +
    scale_y_continuous(expand = c(0, 0)) +
    theme_classic(base_size = 16) +
    theme(
      legend.position  = "bottom",
      legend.title     = element_blank(),
      legend.text      = element_text(size = 14),
      axis.title       = element_text(face = "bold", size = 16),
      axis.text        = element_text(size = 14),
      plot.title       = element_text(face = "bold", size = 18, hjust = 0.5),
      axis.line        = element_line(linewidth = 0.8),
      panel.border     = element_rect(color = "black", fill = NA, linewidth = 0.8)
    ) +
    labs(
      x     = "Year",
      y     = "Predicted Yield (MT/ha)",
      title = paste0(crop_name, " — Bootstrap Trajectories (n = 5,000)")
    )
  
  ggsave(file_name, p, width = 8, height = 4.5, dpi = 300)
  p
}

corn_traj_combined <- bind_rows(
  all_trajectories$corn_irrig %>% mutate(Type = "Irrigated"),
  all_trajectories$corn_rain  %>% mutate(Type = "Rainfed")
) %>%
  left_join(co2_ssp245 %>% select(year, co2_mult_maize),
            by = c("cip_year" = "year")) %>%
  mutate(mean_yield = mean_yield * co2_mult_maize) %>%
  select(-co2_mult_maize)

soy_traj_combined <- bind_rows(
  all_trajectories$soy_irrig %>% mutate(Type = "Irrigated"),
  all_trajectories$soy_rain  %>% mutate(Type = "Rainfed")
) %>%
  left_join(co2_ssp245 %>% select(year, co2_mult_soy),
            by = c("cip_year" = "year")) %>%
  mutate(mean_yield = mean_yield * co2_mult_soy) %>%
  select(-co2_mult_soy)

# Plot — these may take a minute to render with 5000 lines each
corn_spaghetti <- make_spaghetti_plot(
  corn_traj_combined, corn_summary,
  "Maize", "output_data/corn_spaghetti.png"
)

soy_spaghetti <- make_spaghetti_plot(
  soy_traj_combined, soy_summary,
  "Soy", "output_data/soy_spaghetti.png"
)

corn_spaghetti
soy_spaghetti
```
```{r yield-spaghetti-plot-publication}
library(patchwork)

make_spaghetti_plot <- function(traj_df, summary_df, crop_name, panel_letter) {
  
  cols <- c("Irrigated" = "#1f78b4", "Rainfed" = "#33a02c")
  
  ggplot() +
    geom_line(
      data = traj_df,
      aes(x = cip_year, y = mean_yield, 
          group = interaction(boot_iter, gcm, Type),
          color = Type),
      alpha = 0.005,
      linewidth = 0.3
    ) +
    geom_line(
      data = summary_df,
      aes(x = year, y = mean_yield,
          color = Type, group = Type),
      linewidth = 1.8
    ) +
    scale_color_manual(
      values = cols,
      labels = c("Irrigated" = "IRRIGATED", "Rainfed" = "RAINFED")
    ) +
    scale_x_continuous(
      expand = expansion(mult = c(0, 0.02)),
      breaks = c(2015, 2050, 2100),
      limits = c(2015, 2100)
    ) +
    scale_y_continuous(expand = c(0, 0)) +
    theme_classic(base_size = 8) +
    theme(
      legend.position  = "bottom",
      legend.title     = element_blank(),
      legend.text      = element_text(size = 8, family = "Helvetica"),
      axis.title       = element_text(size = 8, family = "Helvetica"),
      axis.text        = element_text(size = 7, family = "Helvetica"),
      plot.title       = element_text(size = 8, face = "bold",
                                      hjust = 0, family = "Helvetica"),
      axis.line        = element_line(linewidth = 0.5),
      panel.border     = element_rect(color = "black", fill = NA,
                                      linewidth = 0.5),
      plot.margin      = margin(4, 8, 4, 4, "pt")
    ) +
    labs(
      x     = "Year",
      y     = "Predicted Yield (MT/ha)",
      title = paste0("(", panel_letter, ") ", crop_name)
    )
}

# Prepare trajectory data frames with CO2 adjustment only (already in MT/ha)
corn_traj_combined <- bind_rows(
  all_trajectories$corn_irrig %>% mutate(Type = "Irrigated"),
  all_trajectories$corn_rain  %>% mutate(Type = "Rainfed")
) %>%
  left_join(co2_ssp245 %>% select(year, co2_mult_maize),
            by = c("cip_year" = "year")) %>%
  mutate(mean_yield = mean_yield * co2_mult_maize) %>%
  select(-co2_mult_maize)

soy_traj_combined <- bind_rows(
  all_trajectories$soy_irrig %>% mutate(Type = "Irrigated"),
  all_trajectories$soy_rain  %>% mutate(Type = "Rainfed")
) %>%
  left_join(co2_ssp245 %>% select(year, co2_mult_soy),
            by = c("cip_year" = "year")) %>%
  mutate(mean_yield = mean_yield * co2_mult_soy) %>%
  select(-co2_mult_soy)

# Build individual panels
p_maize <- make_spaghetti_plot(corn_traj_combined, corn_summary, "Maize", "a")
p_soy   <- make_spaghetti_plot(soy_traj_combined,  soy_summary,  "Soy",   "b")

# Apply shared legend theme to each panel individually before combining
p_maize <- p_maize + theme(legend.position = "bottom")
p_soy   <- p_soy   + theme(legend.position = "bottom")

# Combine and save as single figure
p_combined <- (p_maize / p_soy) +
  plot_layout(guides = "collect")

ggsave(
  "output_data/spaghetti_combined.png",
  p_combined,
  width  = 7.5,
  height = 6,
  dpi    = 600,
  units  = "in"
)

p_combined
```