---
title: "CMIP6 Projected Crop Yield Analysis (SSP2-4.5)"
author: "Penny Hunter"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  warning = FALSE,
  message = FALSE
)
library(dplyr)
library(lubridate)
library(ggplot2)
library(arrow)
```

---

## Overview

This script applies historically-calibrated yield regression models to
bias-corrected CMIP6 projections (SSP2-4.5) to estimate future maize and
soybean yields across US Corn Belt counties through 2100. Five GCMs are used
to represent model uncertainty.

Pipeline:

1. Load training data and fitted regression models
2. Load and crop CMIP6 parquet files to analysis counties
3. Compute daily GDD and KDD from projected temperatures
4. Aggregate to county-year summaries
5. Predict yields and summarize across GCMs
6. Plot projections with GCM uncertainty envelope
7. Compute start-to-end yield changes

---

## 1. Load Training Data and Fitted Models

County factor levels must match those used during model training.

```{r load-data-models}
maize_gs    <- readRDS("input_data/maize_gs")
soybeans_gs <- readRDS("input_data/soybeans_gs")

maize_gs <- maize_gs %>%
  mutate(county_factor = factor(fips))

soybeans_gs <- soybeans_gs %>%
  mutate(county_factor = factor(fips))

model_soy_irrig   <- readRDS("input_data/model_soy_irrig.rds")
model_soy_rain    <- readRDS("input_data/model_soy_rain.rds")
model_maize_irrig <- readRDS("input_data/model_maize_irrig.rds")
model_maize_rain  <- readRDS("input_data/model_maize_rain.rds")

maize_counties   <- as.character(unique(maize_gs$fips))
soybean_counties <- as.character(unique(soybeans_gs$fips))
```

> **Note:** To refit models from scratch, see `prism_data_organizing.Rmd`.
> The code block below is provided for reference and is not evaluated by default.

```{r fit-models, eval=FALSE}
formula_irrig <- irr ~ KDD_sum_yearly + GDD_sum_yearly +
                       prism_ppt + I(prism_ppt^2) +
                       county_factor + factor(year)

formula_rain  <- rainfed ~ KDD_sum_yearly + GDD_sum_yearly +
                           prism_ppt + I(prism_ppt^2) +
                           county_factor + factor(year)

model_soy_irrig   <- lm(formula_irrig, data = soybeans_gs)
model_soy_rain    <- lm(formula_rain,  data = soybeans_gs)
model_maize_irrig <- lm(formula_irrig, data = maize_gs)
model_maize_rain  <- lm(formula_rain,  data = maize_gs)

saveRDS(model_soy_irrig,   "input_data/model_soy_irrig.rds")
saveRDS(model_soy_rain,    "input_data/model_soy_rain.rds")
saveRDS(model_maize_irrig, "input_data/model_maize_irrig.rds")
saveRDS(model_maize_rain,  "input_data/model_maize_rain.rds")
```

---

## 2. Load and Crop CMIP6 Data

Each GCM parquet file is filtered to the counties present in the historical
yield analysis. Only variables required for degree-day calculations are
retained.

```{r crop-cmip-function}
crop_cmip_to_training <- function(parquet_path, counties) {
  vars_needed <- c("GEOID", "time", "tasmax", "tasmin",
                   "pr", "daily_mean_vpd_tdew_bc_temp")
  open_dataset(parquet_path) |>
    select(any_of(vars_needed)) |>
    filter(GEOID %in% counties) |>
    collect()
}
```

```{r load-cmip, cache=TRUE}
cmip_files <- list(
  GFDL  = "input_data/all_GFDL-ESM4_ssp245_bc.parquet",
  IPSL  = "input_data/all_IPSL-CM6A-LR_ssp245_bc.parquet",
  MPI   = "input_data/all_MPI-ESM1-2-HR_ssp245_bc.parquet",
  MRI   = "input_data/all_MRI-ESM2-0_ssp245_bc.parquet",
  UKESM = "input_data/all_UKESM1-0-LL_ssp245_bc.parquet"
)

cmip_cropped <- list()

for (m in names(cmip_files)) {
  message("Loading GCM: ", m)
  cmip_cropped[[m]] <- list(
    maize = crop_cmip_to_training(cmip_files[[m]], maize_counties),
    soy   = crop_cmip_to_training(cmip_files[[m]], soybean_counties)
  )
  # Ensure 5-digit GEOID formatting
  cmip_cropped[[m]]$maize <- cmip_cropped[[m]]$maize %>%
    mutate(GEOID = sprintf("%05s", GEOID))
  cmip_cropped[[m]]$soy <- cmip_cropped[[m]]$soy %>%
    mutate(GEOID = sprintf("%05s", GEOID))
}
```

---

## 3. Compute Daily GDD and KDD

Growing Degree Days (GDD) and Killing Degree Days (KDD) use the same hourly
temperature profile as the historical analysis (Parton & Logan 1981).
Vectorised matrix arithmetic is used for efficiency.

```{r degree-day-functions}
kdd_threshold <- 30
gdd_base      <- 10
gdd_cap       <- 30

Tf <- 0.44 - 0.46 * sin((pi * seq(0.5, 23, 1)) / 12 + 0.9) +
      0.11 * sin((2 * pi * seq(0.5, 23, 1)) / 12 + 0.9)

compute_kdd_gdd <- function(df) {
  hrlyT <- outer(df$tasmax, Tf) + outer(df$tasmin, (1 - Tf))

  # KDD: hours above heat stress threshold
  df$KDD_daily <- rowMeans(pmax(hrlyT - kdd_threshold, 0))

  # GDD: hours within biological temperature range
  hrlyT_GDD    <- pmin(pmax(hrlyT, gdd_base), gdd_cap)
  df$GDD_daily <- rowMeans(pmax(hrlyT_GDD - gdd_base, 0))

  df
}
```

---

## 4. Aggregate to County-Year Summaries

```{r aggregate-function}
aggregate_county_year <- function(df) {
  df %>%
    mutate(year = year(time)) %>%
    group_by(GEOID, year) %>%
    summarise(
      GDD_sum_yearly = sum(GDD_daily, na.rm = TRUE),
      KDD_sum_yearly = sum(KDD_daily, na.rm = TRUE),
      prism_ppt      = sum(pr,        na.rm = TRUE),
      vpd_mean       = mean(daily_mean_vpd_tdew_bc_temp, na.rm = TRUE),
      .groups        = "drop"
    )
}
```

---

## 5. Prepare Data for Prediction

County factor levels are aligned to the training data. The technology trend
is held constant at 2018 (the final historical year) to isolate the climate
signal.

```{r prediction-functions}
frozen_year <- 2018

prepare_for_prediction <- function(df, model_object) {
  df <- df %>%
    mutate(GEOID = sprintf("%05s", GEOID)) %>%
    filter(GEOID %in% levels(model_object$model$county_factor)) %>%
    mutate(county_factor = factor(GEOID,
                                  levels = levels(model_object$model$county_factor)))

  required_cols <- c("year", "GDD_sum_yearly", "KDD_sum_yearly",
                     "prism_ppt", "vpd_mean", "cip_year", "county_factor")
  missing_cols <- setdiff(required_cols, names(df))
  if (length(missing_cols) > 0) {
    stop("Missing columns: ", paste(missing_cols, collapse = ", "))
  }

  df$predicted_yield <- predict(model_object, newdata = df)
  df
}

process_crop_for_model <- function(cropped_df, model_object, frozen_year) {
  cropped_df %>%
    compute_kdd_gdd() %>%
    aggregate_county_year() %>%
    mutate(
      cip_year = year,
      year     = frozen_year
    ) %>%
    prepare_for_prediction(model_object)
}
```

---

## 6. Run Projections Across All GCMs

```{r run-projections, cache=TRUE}
maize_models <- list(irrigated = model_maize_irrig, rainfed = model_maize_rain)
soy_models   <- list(irrigated = model_soy_irrig,   rainfed = model_soy_rain)

results <- list()

for (gcm in names(cmip_cropped)) {
  message("Processing GCM: ", gcm)
  results[[gcm]] <- list(
    maize = list(
      irrigated = process_crop_for_model(cmip_cropped[[gcm]]$maize,
                                         maize_models$irrigated, frozen_year),
      rainfed   = process_crop_for_model(cmip_cropped[[gcm]]$maize,
                                         maize_models$rainfed,   frozen_year)
    ),
    soy = list(
      irrigated = process_crop_for_model(cmip_cropped[[gcm]]$soy,
                                         soy_models$irrigated, frozen_year),
      rainfed   = process_crop_for_model(cmip_cropped[[gcm]]$soy,
                                         soy_models$rainfed,   frozen_year)
    )
  )
}
```

---

## 7. Summarize Across GCMs

Mean yield per year is computed within each GCM, then the ensemble mean,
minimum, and maximum are derived across GCMs to represent projection
uncertainty.

```{r summarize-gcms}
summarize_yield_gcms <- function(gcm_list, model) {
  gcm_means <- lapply(seq_along(gcm_list), function(i) {
    df   <- gcm_list[[i]]
    pred <- predict(model, newdata = df)
    df %>%
      mutate(predicted_yield = pmax(pred, 0)) %>%
      group_by(cip_year) %>%
      summarise(mean_yield_gcm = mean(predicted_yield, na.rm = TRUE),
                .groups = "drop") %>%
      mutate(GCM = names(gcm_list)[i])
  })

  bind_rows(gcm_means) %>%
    group_by(cip_year) %>%
    summarise(
      mean_yield = mean(mean_yield_gcm, na.rm = TRUE),
      min_yield  = min(mean_yield_gcm,  na.rm = TRUE),
      max_yield  = max(mean_yield_gcm,  na.rm = TRUE),
      .groups    = "drop"
    ) %>%
    filter(cip_year != 2014) %>%
    rename(year = cip_year)
}

corn_df <- bind_rows(
  summarize_yield_gcms(lapply(results, \(x) x$maize$irrigated), model_maize_irrig) %>% mutate(Type = "Irrigated"),
  summarize_yield_gcms(lapply(results, \(x) x$maize$rainfed),   model_maize_rain)  %>% mutate(Type = "Rainfed")
)

soy_df <- bind_rows(
  summarize_yield_gcms(lapply(results, \(x) x$soy$irrigated), model_soy_irrig) %>% mutate(Type = "Irrigated"),
  summarize_yield_gcms(lapply(results, \(x) x$soy$rainfed),   model_soy_rain)  %>% mutate(Type = "Rainfed")
)
```

---

## 8. Plot Projected Yields

Shaded ribbons represent the GCM ensemble range (min–max). Lines show the
ensemble mean.

```{r plot-function}
cols <- c("Irrigated" = "#1f78b4", "Rainfed" = "#33a02c")

make_plot_minmax <- function(df, crop_name, file_name) {
  p <- ggplot(df, aes(x = year, y = mean_yield, color = Type, fill = Type)) +
    geom_ribbon(aes(ymin = min_yield, ymax = max_yield, group = Type),
                alpha = 0.2, color = NA) +
    geom_line(aes(group = Type), linewidth = 1.2) +
    scale_color_manual(values = cols,
                       labels = c("Irrigated" = "IRRIGATED", "Rainfed" = "RAINFED")) +
    scale_fill_manual(values  = cols,
                      labels  = c("Irrigated" = "IRRIGATED", "Rainfed" = "RAINFED")) +
    scale_x_continuous(expand = c(0, 0),
                       breaks = c(min(df$year), 2050, 2100)) +
    scale_y_continuous(expand = c(0, 0)) +
    labs(x = "Year", y = "Predicted Yield (bu/ac)", title = crop_name) +
    theme_classic(base_size = 16) +
    theme(
      legend.position = "bottom",
      legend.title    = element_blank(),
      legend.text     = element_text(size = 14),
      axis.title      = element_text(face = "bold", size = 16),
      axis.text       = element_text(size = 14),
      plot.title      = element_text(face = "bold", size = 18, hjust = 0.5),
      axis.line       = element_line(linewidth = 0.8),
      panel.border    = element_rect(color = "black", fill = NA, linewidth = 0.8)
    )

  ggsave(file_name, p, width = 8, height = 4.5, dpi = 300)
  p
}
```

```{r save-plots, fig.width=8, fig.height=4.5}
make_plot_minmax(corn_df, "Corn", "output_data/corn_plot_minmax.png")
make_plot_minmax(soy_df,  "Soy",  "output_data/soy_plot_minmax.png")
```

---

## 9. Compute Start-to-End Yield Changes

Absolute and percentage yield change from the first to the last projected year,
summarized by crop and irrigation type.

```{r yield-changes}
compute_yield_change <- function(df, crop, type_label) {
  df %>%
    filter(Type == type_label) %>%
    arrange(year) %>%
    summarise(
      Crop        = crop,
      Type        = type_label,
      Start_Year  = first(year),
      End_Year    = last(year),
      Start_Yield = round(first(mean_yield), 2),
      End_Yield   = round(last(mean_yield),  2),
      Abs_Change  = round(last(mean_yield) - first(mean_yield), 2),
      Pct_Change  = round(((last(mean_yield) - first(mean_yield)) / first(mean_yield)) * 100, 1)
    )
}

yield_changes <- bind_rows(
  compute_yield_change(corn_df, "Corn", "Irrigated"),
  compute_yield_change(corn_df, "Corn", "Rainfed"),
  compute_yield_change(soy_df,  "Soy",  "Irrigated"),
  compute_yield_change(soy_df,  "Soy",  "Rainfed")
)

knitr::kable(yield_changes,
             caption = "Projected yield change from start to end of simulation period.")
```
