---
title: "PRISM Climate Data Processing & Crop Yield Regression Analysis"
author: "Penny Hunter"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  warning = FALSE,
  message = FALSE
)
source("~/Desktop/ASCI-GitHub/Masters-Thesis/setup.R")
```

---

## Overview

This script processes daily PRISM climate data and USDA crop yield records to
construct county-level growing-season datasets for maize and soybeans across
the US Corn Belt. The pipeline covers:

1. Loading and merging PRISM climate variables
2. Extracting crop-specific growing seasons from USDA Crop Progress data
3. Calculating daily Growing Degree Days (GDD) and Killing Degree Days (KDD)
4. Aggregating to annual county-level summaries
5. Merging with USDA yield data and applying data quality filters
6. Running panel regressions and state-year block bootstrap inference

---

## 1. Load PRISM Climate Data

```{r load-prism}
ppt_county_full    <- readRDS("input_data/prism_data_full/ppt_county_full.rds")
tmin_county_full   <- readRDS("input_data/prism_data_full/tmin_county_full.rds")
tmax_county_full   <- readRDS("input_data/prism_data_full/tmax_county_full.rds")
vpdmin_county_full <- readRDS("input_data/prism_data_full/vpdmin_county_full.rds")
vpdmean_county_full<- readRDS("input_data/prism_data_full/vpdmean_county_full.rds")
vpdmax_county_full <- readRDS("input_data/prism_data_full/vpdmax_county_full.rds")
```

---

## 2. Prepare PRISM Datasets for Merging

Drop redundant identifier columns before joining across variables.

```{r prep-prism}
prism_datasets <- list(
  tmin_county_full,
  ppt_county_full,
  vpdmax_county_full,
  vpdmin_county_full,
  vpdmean_county_full
)

dataset_names <- c(
  "tmin_county_full",
  "ppt_county_full",
  "vpdmax_county_full",
  "vpdmin_county_full",
  "vpdmean_county_full"
)

cols_to_remove <- c("county_name", "county_count", "Year", "Month", "Day")

for (i in seq_along(prism_datasets)) {
  df <- prism_datasets[[i]]
  df <- df[, !(names(df) %in% cols_to_remove)]
  assign(dataset_names[i], df, envir = .GlobalEnv)
}
```

---

## 3. Merge All PRISM Variables

Join all climate variables into a single panel dataset keyed by county (GEOID)
and date (time).

```{r merge-prism}
merged_prism_full <- tmax_county_full %>%
  full_join(tmin_county_full,    by = c("time", "GEOID")) %>%
  full_join(ppt_county_full,     by = c("time", "GEOID")) %>%
  full_join(vpdmax_county_full,  by = c("time", "GEOID")) %>%
  full_join(vpdmean_county_full, by = c("time", "GEOID")) %>%
  full_join(vpdmin_county_full,  by = c("time", "GEOID"))

merged_prism_full_1 <- merged_prism_full

saveRDS(
  merged_prism_full_1,
  "/Users/pennyhunter/Desktop/ASCI-GitHub/Masters-Thesis/input_data/merged_prism_full_1.RDS"
)

rm(merged_prism_full, ppt_county_full, tmin_county_full, tmax_county_full,
   vpdmin_county_full, vpdmean_county_full, vpdmax_county_full,
   prism_datasets, df)
```

---

## 4. Load Growing Season Dates from USDA Crop Progress

Median emergence and harvest day-of-year by state, derived from USDA Crop
Progress reports.

```{r load-growing-seasons}
maize_median    <- read_csv("input_data/maize_median.csv")
soybeans_median <- read_csv("input_data/soybeans_median.csv")

condense_seasons <- function(df) {
  df %>%
    select(state_name, state_id, growing_phase, start_doy) %>%
    filter(growing_phase %in% c("EMERGED", "HARVESTED")) %>%
    pivot_wider(names_from = growing_phase, values_from = start_doy) %>%
    rename(emerged = EMERGED, harvested = HARVESTED)
}

maize_condensed    <- condense_seasons(maize_median)
soybeans_condensed <- condense_seasons(soybeans_median)

rm(maize_median, soybeans_median)
```

---

## 5. Subset PRISM Data to Crop Growing Seasons

Retain only observations falling within each crop's state-specific growing
season window.

```{r subset-growing-season}
merged_prism_full_1 <- merged_prism_full_1 %>%
  mutate(
    time     = as.Date(time),
    doy      = yday(time),
    state_id = as.integer(str_sub(GEOID, 1, 2))
  )

maize_gs <- merged_prism_full_1 %>%
  left_join(maize_condensed, by = "state_id") %>%
  filter(doy >= emerged & doy <= harvested) %>%
  select(-emerged, -harvested)

soybeans_gs <- merged_prism_full_1 %>%
  left_join(soybeans_condensed, by = "state_id") %>%
  filter(doy >= emerged & doy <= harvested) %>%
  select(-emerged, -harvested)

rm(merged_prism_full_1, maize_condensed, soybeans_condensed)
```

---

## 6. Construct Daily GDD and KDD

Growing Degree Days (GDD) and Killing Degree Days (KDD) are calculated using
an hourly temperature profile derived from daily Tmax and Tmin following
Schlenker & Roberts (2009).

```{r gdd-kdd-functions}
kdd_threshold <- 30  # Heat stress threshold (°C)
gdd_base      <- 10  # Biological base temperature (°C)
gdd_cap       <- 30  # Biological ceiling temperature (°C)

# Hourly temperature fraction profile (Parton & Logan 1981)
Tf <- 0.44 - 0.46 * sin((pi * seq(0.5, 23, 1)) / 12 + 0.9) +
      0.11 * sin((2 * pi * seq(0.5, 23, 1)) / 12 + 0.9)

calculate_kdd <- function(T_max, T_min, thr, Tf) {
  hrlyT <- T_max * Tf + T_min * (1 - Tf)
  sum(pmax(hrlyT - thr, 0)) / 24
}

calculate_gdd <- function(T_max, T_min, Tf, base = 10, cap = 30) {
  T_max <- min(T_max, cap)
  T_min <- max(T_min, base)
  hrlyT <- T_max * Tf + T_min * (1 - Tf)
  sum(pmax(hrlyT - base, 0)) / 24
}
```

```{r apply-gdd-kdd}
apply_degree_days <- function(df) {
  df$KDD <- mapply(calculate_kdd,
                   T_max = df$prism_tmax, T_min = df$prism_tmin,
                   MoreArgs = list(thr = kdd_threshold, Tf = Tf))
  df$GDD <- mapply(calculate_gdd,
                   T_max = df$prism_tmax, T_min = df$prism_tmin,
                   MoreArgs = list(Tf = Tf, base = gdd_base, cap = gdd_cap))
  df
}

maize_gs    <- apply_degree_days(maize_gs)
soybeans_gs <- apply_degree_days(soybeans_gs)
```

---

## 7. Aggregate to Annual County-Level Summaries

Sum KDD and GDD over the growing season; average remaining climate variables
by county-year.

```{r annual-aggregation}
aggregate_annual <- function(df) {
  df %>%
    mutate(year = year(time)) %>%
    group_by(GEOID, year) %>%
    mutate(
      KDD_sum_yearly = sum(KDD, na.rm = TRUE),
      GDD_sum_yearly = sum(GDD, na.rm = TRUE)
    ) %>%
    ungroup() %>%
    select(-KDD, -GDD) %>%
    group_by(GEOID, year) %>%
    summarise(
      county_count   = first(county_count),
      prism_tmax     = mean(prism_tmax,    na.rm = TRUE),
      prism_tmin     = mean(prism_tmin,    na.rm = TRUE),
      prism_vpdmean  = mean(prism_vpdmean, na.rm = TRUE),
      prism_vpdmax   = mean(prism_vpdmax,  na.rm = TRUE),
      prism_vpdmin   = mean(prism_vpdmin,  na.rm = TRUE),
      prism_ppt      = sum(prism_ppt,      na.rm = TRUE),
      KDD_sum_yearly = first(KDD_sum_yearly),
      GDD_sum_yearly = first(GDD_sum_yearly),
      .groups = "drop"
    )
}

maize_gs    <- aggregate_annual(maize_gs)
soybeans_gs <- aggregate_annual(soybeans_gs)
```

---

## 8. Merge with USDA Yield Data

```{r merge-yield}
yield_benefit_model_data <- read_csv("input_data/yield-benefit-model-data.csv")

crop_datasets <- split(yield_benefit_model_data, yield_benefit_model_data$commodity_desc)
list2env(crop_datasets, envir = .GlobalEnv)

CORN <- CORN %>% filter(util_practice_desc == "GRAIN")

rm(OATS, PEANUTS, SORGHUM, WHEAT, HAY, COTTON, BEANS, BARLEY,
   yield_benefit_model_data, crop_datasets)

maize_gs <- merge(CORN,     maize_gs,    by.x = c("fips", "year"), by.y = c("GEOID", "year"), all.x = TRUE)
soybeans_gs <- merge(SOYBEANS, soybeans_gs, by.x = c("fips", "year"), by.y = c("GEOID", "year"), all.x = TRUE)

rm(CORN, SOYBEANS)
```

---

## 9. Apply Data Quality Filters

Retain only counties with at least 10 years of yield observations.

```{r quality-filter}
maize_gs <- maize_gs %>%
  group_by(fips) %>%
  filter(n_distinct(year) >= 10) %>%
  ungroup()

soybeans_gs <- soybeans_gs %>%
  group_by(fips) %>%
  filter(n_distinct(year) >= 10) %>%
  ungroup()
```

---

## 10. Finalize and Save Processed Datasets

```{r save-datasets}
# Save county crosswalk for downstream spatial work
counties_combined <- bind_rows(
  maize_gs    %>% select(fips, county_ansi, state_ansi),
  soybeans_gs %>% select(fips, county_ansi, state_ansi)
) %>% distinct()

saveRDS(counties_combined, file = "input_data/counties_combined.rds")

# Create factor variables for panel regression
maize_gs <- maize_gs %>%
  mutate(
    county_factor = as.factor(county_ansi),
    state_factor  = as.factor(state_ansi)
  )

soybeans_gs <- soybeans_gs %>%
  mutate(
    county_factor = as.factor(county_ansi),
    state_factor  = as.factor(state_ansi)
  )

# Convert precipitation to cm (from mm)
maize_gs    <- maize_gs    %>% mutate(prism_ppt_cm = prism_ppt / 10)
soybeans_gs <- soybeans_gs %>% mutate(prism_ppt_cm = prism_ppt / 10)

saveRDS(maize_gs,    file = "input_data/maize_gs")
saveRDS(soybeans_gs, file = "input_data/soybeans_gs")

crops <- list(corn = maize_gs, soybeans = soybeans_gs)
```

---
## 11. State-Year Block Bootstrap: KDD Sensitivity

Bootstrap confidence intervals for the KDD coefficient using state-year block
resampling to account for spatial and temporal dependence.

```{r bootstrap-function}
set.seed(123)
n_boot <- 1000

bootstrap_lm_state_year <- function(data, formula, n_boot = 1000,
                                    state_var = "state_ansi", year_var = "year") {
  base_model <- lm(formula, data = data)
  coef_names <- names(coef(base_model))

  data <- data %>%
    mutate(state_year = interaction(.data[[state_var]], .data[[year_var]], drop = TRUE))
  unique_blocks <- unique(data$state_year)

  boot_coefs <- replicate(n_boot, {
    sampled_blocks <- sample(unique_blocks, size = length(unique_blocks), replace = TRUE)
    boot_sample    <- do.call(rbind, lapply(sampled_blocks, function(b) data[data$state_year == b, ]))

    tryCatch({
      coefs      <- coef(lm(formula, data = boot_sample))
      coefs_full <- setNames(rep(NA, length(coef_names)), coef_names)
      coefs_full[intersect(names(coefs), coef_names)] <- coefs[intersect(names(coefs), coef_names)]
      coefs_full
    }, error = function(e) setNames(rep(NA, length(coef_names)), coef_names))
  }, simplify = TRUE)

  as.data.frame(t(boot_coefs))
}
```

```{r run-bootstrap}
boot_results <- list()

for (crop_name in names(crops)) {
  crop_data <- crops[[crop_name]]

  for (water_type in c("rainfed", "irr")) {
    if (!water_type %in% names(crop_data)) next

    formula <- as.formula(paste0(
      water_type,
      " ~ KDD_sum_yearly + GDD_sum_yearly + prism_ppt_cm + I(prism_ppt_cm^2) +",
      " county_factor + factor(year)"
    ))

    boot_df   <- bootstrap_lm_state_year(crop_data, formula, n_boot)
    kdd_coefs <- na.omit(boot_df$KDD_sum_yearly)
    if (length(kdd_coefs) < 10) next

    boot_results[[length(boot_results) + 1]] <- data.frame(
      Crop  = tools::toTitleCase(crop_name),
      Type  = toupper(ifelse(water_type == "irr", "IRRIGATED", "RAINFED")),
      Mean  = mean(kdd_coefs),
      Lower = quantile(kdd_coefs, 0.025),
      Upper = quantile(kdd_coefs, 0.975)
    )
  }
}

boot_results_df <- bind_rows(boot_results)
```

```{r plot-bootstrap, fig.width=9, fig.height=6.5}
nice_colors <- c("IRRIGATED" = "#1f78b4", "RAINFED" = "#33a02c")

kdd_plot <- ggplot(boot_results_df, aes(x = Crop, y = Mean, color = Type)) +
  geom_point(size = 5.4, position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = Lower, ymax = Upper),
                width = 0.2, linewidth = 2.7,
                position = position_dodge(width = 0.5)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50", linewidth = 0.9) +
  scale_color_manual(values = nice_colors) +
  labs(
    x     = expression(bold("Crop Type")),
    y     = expression(bold(beta[1] ~ "(KDD)")),
    color = NULL
  ) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major  = element_blank(),
    panel.grid.minor  = element_blank(),
    panel.border      = element_rect(color = "black", fill = NA, linewidth = 1),
    axis.title        = element_text(size = 20, face = "bold"),
    axis.text         = element_text(size = 16),
    axis.text.x       = element_text(angle = 45, hjust = 1),
    legend.position   = "right",
    legend.key.size   = unit(1.6, "lines"),
    legend.text       = element_text(size = 16),
    legend.spacing.y  = unit(0.35, "cm")
  )

kdd_plot

ggsave(
  filename = "output_data/KDD_bootstrap_coefficients.png",
  plot     = kdd_plot,
  width    = 9, height = 6.5, dpi = 600
)
```

---

## 12. Export Regression Tables to Word

Coefficient tables with bootstrap confidence intervals exported for each
crop-by-irrigation combination.

```{r export-word-tables}
library(officer)

coef_keep <- c(
  "KDD_sum_yearly",
  "GDD_sum_yearly",
  "prism_ppt_cm",
  "I(prism_ppt_cm^2)"
)

pretty_labels <- c(
  KDD_sum_yearly       = "Killing Degree Days",
  GDD_sum_yearly       = "Growing Degree Days",
  prism_ppt_cm         = "Precipitation (cm)",
  "I(prism_ppt_cm^2)"  = "Precipitation\u00B2 (cm)"
)

get_stars <- function(ci_low, ci_high) {
  if (ci_low > 0 | ci_high < 0) {
    if (abs(ci_low) > 0.1 & abs(ci_high) > 0.1) "***"
    else if (abs(ci_low) > 0.05 & abs(ci_high) > 0.05) "**"
    else "*"
  } else ""
}

for (crop_name in names(crops)) {
  crop_data  <- crops[[crop_name]]
  crop_title <- tools::toTitleCase(crop_name)

  for (water_type in c("rainfed", "irr")) {
    if (!water_type %in% names(crop_data)) next

    water_title <- ifelse(water_type == "irr", "Irrigated", "Rainfed")

    formula <- as.formula(paste0(
      water_type,
      " ~ KDD_sum_yearly + GDD_sum_yearly + prism_ppt_cm + I(prism_ppt_cm^2) +",
      " county_factor + factor(year)"
    ))

    model   <- lm(formula, data = crop_data)
    boot_df <- bootstrap_lm_state_year(crop_data, formula, n_boot,
                                       state_var = "state_ansi", year_var = "year")

    coef_rows <- lapply(coef_keep, function(term) {
      if (!term %in% names(boot_df)) return(NULL)
      x <- na.omit(boot_df[[term]])
      if (length(x) < 10) return(NULL)

      ci_low  <- quantile(x, 0.025)
      ci_high <- quantile(x, 0.975)

      data.frame(
        Predictor = pretty_labels[term],
        Estimate  = paste0(round(coef(model)[term], 3), get_stars(ci_low, ci_high)),
        `95% CI`  = paste0("[", round(ci_low, 3), ", ", round(ci_high, 3), "]"),
        stringsAsFactors = FALSE
      )
    })

    table_df <- bind_rows(coef_rows)

    stats_df <- data.frame(
      Predictor = c("Observations", "R-squared"),
      Estimate  = as.character(c(nobs(model), round(summary(model)$r.squared, 3))),
      `95% CI`  = c("", ""),
      stringsAsFactors = FALSE
    )

    table_df <- bind_rows(table_df, stats_df)

    doc <- read_docx() %>%
      body_add_par(paste0(crop_title, " \u2014 ", water_title), style = "heading 1") %>%
      body_add_table(value = table_df) %>%
      body_add_par(
        paste0("Notes: County and year fixed effects included. ",
               "95% CIs based on state-year block bootstrap (", n_boot, " replications). ",
               "* p<0.05, ** p<0.01, *** p<0.001."),
        style = "Normal"
      )

    print(doc, target = paste0(crop_name, "_", water_title, ".docx"))
  }
}
```
