---
title: "Prism Data Organizing"
author: "Penelope Hunter"
date: "2025-03-12"
output: html_document
---

#0.0 Setup packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("~/Desktop/ASCI-GitHub/Masters-Thesis/setup.R")
```


#1.0 PRISM Climate Variable Merging 
#The following section creates a dataset of PRISM Climate Variables if your computer can handle loading in the whole dataset please skip to optional section 2.0

#1.1 Loading Data

```{r}
# Define the folder path
data_folder <- "input_data/LDAS_SM_0_10"

# List all CSV files in the folder
all_files <- list.files(path = data_folder, pattern = "\\.csv$", full.names = TRUE)

# Filter files for years <= 2018
files_to_load <- all_files %>%
  grep(pattern = "_19[8-9][0-9]\\.csv$|_200[0-9]\\.csv$|_201[0-8]\\.csv$", value = TRUE)

# Read and combine all selected files into one data frame
ldas_soil_moisture <- lapply(files_to_load, read.csv) %>%
  bind_rows()

```


```{r}
#Loading the climate variable data from PRISM data 
ppt_county_full <- readRDS("input_data/prism_data_full/ppt_county_full.rds")
tmin_county_full <- readRDS("input_data/prism_data_full/tmin_county_full.rds")
tmax_county_full <- readRDS("input_data/prism_data_full/tmax_county_full.rds")
vpdmin_county_full <- readRDS("input_data/prism_data_full/vpdmin_county_full.rds")
vpdmean_county_full <- readRDS("input_data/prism_data_full/vpdmean_county_full.rds")
vpdmax_county_full <- readRDS("input_data/prism_data_full/vpdmax_county_full.rds")
```

#1.2 Prep PRISM Datasets for Merge
```{r}
# List all dataset objects you want to clean
prism_datasets <- list(
  tmin_county_full,
  ppt_county_full,
  vpdmax_county_full,
  vpdmin_county_full,
  vpdmean_county_full
)

# Names to assign back
dataset_names <- c(
  "tmin_county_full",
  "ppt_county_full",
  "vpdmax_county_full",
  "vpdmin_county_full",
  "vpdmean_county_full"
)

# Columns to drop
cols_to_remove <- c("county_name", "county_count", "Year", "Month", "Day")

# Apply removal to each dataset
for (i in seq_along(prism_datasets)) {
  df <- prism_datasets[[i]]
  df <- df[, !(names(df) %in% cols_to_remove)]
  assign(dataset_names[i], df, envir = .GlobalEnv)
}

ldas_soil_moisture <- ldas_soil_moisture %>%
  mutate(
    time = as.Date(time),   # Convert time from character to Date
    GEOID = as.character(GEOID),
    SM = var                # Rename var to SM
  ) %>%
  select(GEOID, time, SM)   # Keep only relevant columns    
```

#1.3 Merge PRISM Dataset
```{r}
#Then merge all data sets by time 
#Start with the unchanged tmax data set and use full_join to join others
merged_prism_full <- tmax_county_full %>%
  full_join(tmin_county_full, by = c("time", "GEOID")) %>%
  full_join(ppt_county_full, by = c("time", "GEOID")) %>%
  full_join(vpdmax_county_full, by = c("time", "GEOID")) %>%
  full_join(vpdmean_county_full, by = c("time", "GEOID")) %>%
  full_join(vpdmin_county_full, by = c("time", "GEOID")) %>% 
  full_join(ldas_soil_moisture, by = c("time", "GEOID"))   

merged_prism_full_1 <- merged_prism_full

saveRDS(merged_prism_full_1,"/Users/pennyhunter/Desktop/ASCI-GitHub/Masters-Thesis/input_data/merged_prism_full_1.RDS")
```

#2.0 Optional Loading Merged PRISM Data Set
```{r}
#If your computer has enough memory to load the merged data set directly please do so here 
#merged_prism_full_1 <- readRDS("Masters-Thesis/input_data/merged_prism_full_1.RDS")
```

#Penny I stopped here at 10/29/25 20:44PM
#3.0 Anomalies 
## The following section takes the PRISM Climate Variables and Soil Moisture data and calculates the average anomaly by county for each variable. Then the data is resampled with a boostrap to ...

#3.1 Create moving average and average anomaly for each variable
```{r} 
merged_prism_full_1$moving_avg_vpdmax <- rollapply(merged_prism_full_1$prism_vpdmax, width = 30, FUN = mean, align = 'center', fill = NA)
#now anamoly
merged_prism_full_1$anomaly_vpdmax <- merged_prism_full_1$prism_vpdmax - merged_prism_full_1$moving_avg_vpdmax
#changing mocing average from 15 to 30 days 
merged_prism_full_1$moving_avg_vpdmin <- rollapply(merged_prism_full_1$prism_vpdmin, width = 30, FUN = mean, align = 'center', fill = NA)
#now anamoly
merged_prism_full_1$anomaly_vpdmin <- merged_prism_full_1$prism_vpdmin - merged_prism_full_1$moving_avg_vpdmin
#changing mocing average from 15 to 30 days 
merged_prism_full_1$moving_avg_vpdmean <- rollapply(merged_prism_full_1$prism_vpdmean, width = 30, FUN = mean, align = 'center', fill = NA)
#now anamoly
merged_prism_full_1$anomaly_vpdmean <- merged_prism_full_1$prism_vpdmean - merged_prism_full_1$moving_avg_vpdmean
#changing mocing average from 15 to 30 days 
merged_prism_full_1$moving_avg_ppt <- rollapply(merged_prism_full_1$prism_ppt, width = 30, FUN = mean, align = 'center', fill = NA)
# now anamoly
merged_prism_full_1$anomaly_ppt <- merged_prism_full_1$prism_ppt - merged_prism_full_1$moving_avg_ppt
#changing mocing average from 15 to 30 days 
merged_prism_full_1$moving_avg_tmax <- rollapply(merged_prism_full_1$prism_tmax, width = 30, FUN = mean, align = 'center', fill = NA)
#now anamoly
merged_prism_full_1$anomaly_tmax <- merged_prism_full_1$prism_tmax - merged_prism_full_1$moving_avg_tmax
# --- Soil Moisture Anomaly (formatted same as PRISM) ---
# 30-day moving average of soil moisture (centered)
merged_prism_full_1$moving_avg_SM <- rollapply(merged_prism_full_1$SM, width = 30, FUN = mean, align = "center",fill = NA)
# Anomaly = actual - moving average
merged_prism_full_1$anomaly_SM <- merged_prism_full_1$SM - merged_prism_full_1$moving_avg_SM

```

#3.2 Making the binary hot day variables for later use
```{r}
merged_prism_full_1$binary_30 <- ifelse(!is.na(merged_prism_full_1$prism_tmax) & merged_prism_full_1$prism_tmax >= 30, 1, 0)

merged_prism_full_1$binary_32 <- ifelse(!is.na(merged_prism_full_1$prism_tmax) & merged_prism_full_1$prism_tmax >= 32, 1, 0)
#colnames(merged_prism_full_1)
```

#3.3 Set up the KDD and GDD calculation for the PRISM dataset
```{r}
# --- Set threshold and hourly temperature profile ---
kdd_threshold <- 30  # Heat stress threshold for KDD
gdd_base <- 5        # Base temp for GDD
gdd_cap <- 30        # Max cap temp for GDD

# Vector of hourly temperature fractions (Tf), representing fraction of daily Tmax over 24 hours
Tf <- 0.44 - 0.46 * sin((pi * seq(0.5, 23, 1)) / 12 + 0.9) + 
      0.11 * sin((2 * pi * seq(0.5, 23, 1)) / 12 + 0.9)

# --- KDD function ---
calculate_kdd_vec <- function(T_max, T_min, thr, Tf) {
  hrlyT <- T_max * Tf + T_min * (1 - Tf)
  hourly_KDD <- pmax(hrlyT - thr, 0)
  daily_KDD <- sum(hourly_KDD) / 24
  return(daily_KDD)
}

# --- GDD function ---
calculate_gdd_vec <- function(T_max, T_min, Tf, base = 5, cap = 30) {
  # Exclude extreme days
  if (T_min < base || T_max > cap) {
    return(0)
  }
  hrlyT <- T_max * Tf + T_min * (1 - Tf)
  hourly_GDD <- pmax(hrlyT - base, 0)
  daily_GDD <- sum(hourly_GDD) / 24
  return(daily_GDD)
}

# --- Example dataset (you must already have this) ---
# merged_prism_full_1 <- data.table(
#   prism_tmax = ..., 
#   prism_tmin = ...
# )

# --- Apply KDD calculation ---
merged_prism_full_1$KDD <- mapply(
  calculate_kdd_vec,
  T_max = merged_prism_full_1$prism_tmax,
  T_min = merged_prism_full_1$prism_tmin,
  MoreArgs = list(thr = kdd_threshold, Tf = Tf)
)

# --- Apply GDD calculation ---
merged_prism_full_1$GDD <- mapply(
  calculate_gdd_vec,
  T_max = merged_prism_full_1$prism_tmax,
  T_min = merged_prism_full_1$prism_tmin,
  MoreArgs = list(Tf = Tf, base = gdd_base, cap = gdd_cap)
)

#Adds GDD and KDD by county/year
merged_prism_full_1 <- merged_prism_full_1 %>%
  mutate(year = year(time)) %>%
  group_by(GEOID, year) %>%
  mutate(
    KDD_sum_yearly = sum(KDD, na.rm = TRUE),
    GDD_sum_yearly = sum(GDD, na.rm = TRUE)
  ) %>%
  ungroup()
```


#3.4 Run the boostrapping for each cliamte variables' anomaly
```{r}
# Preparing data frame to data table
setDT(merged_prism_full_1)
merged_prism_full_1[, time := as.Date(time)]

# Filter to Nebraska counties (GEOIDs start with 31)
merged_prism_full_ne <- merged_prism_full_1[GEOID %like% "^31"]

# Build hot_days table with GEOID intact
hot_days <- merged_prism_full_ne[binary_30 == 1, .(
  GEOID,
  day0_date = time,
  KDD
)]
hot_days[, year := year(day0_date)]

# ============================
# Step 2: Define parameters
# ============================
anomaly_vars <- c(
  "anomaly_tmax",
  "anomaly_vpdmax",
  "anomaly_vpdmean",
  "anomaly_vpdmin",
  "anomaly_ppt"
)
day_offsets <- -30:30
n_boot <- 1000
unique_years <- unique(hot_days$year)

set.seed(123)  # for reproducibility

# ============================
# Step 3: Loop through each anomaly variable
# ============================
for (var in anomaly_vars) {
  cat("\nProcessing", var, "\n")

  # --- Build the correct anomaly lookup for this variable ---
  anomaly_lookup <- merged_prism_full_ne[, .(GEOID, time, anomaly_value = get(var))]

  # --- Bootstrap matrix setup ---
  bootstrap_mat <- matrix(NA_real_, nrow = n_boot, ncol = length(day_offsets))
  colnames(bootstrap_mat) <- paste0("d", day_offsets)

  # --- Bootstrap loop ---
  for (i in 1:n_boot) {
    if (i %% 100 == 0) cat("  Bootstrap", i, "of", n_boot, "\n")

    # Sample years with replacement
    sampled_years <- sample(unique_years, replace = TRUE)

    # Select all hot days from sampled years
    hot_sample <- hot_days[year %in% sampled_years]

    # Expand to windows
    boot_windows <- hot_sample[, .(offset = day_offsets), by = .(GEOID, day0_date, KDD)]
    boot_windows[, window_date := day0_date + offset]

    # Join with anomaly values
    boot_window_data <- merge(
      boot_windows,
      anomaly_lookup,
      by.x = c("GEOID", "window_date"),
      by.y = c("GEOID", "time"),
      all.x = TRUE
    )

    # Apply KDD weighting to each row
    boot_window_data[, weighted_anomaly := anomaly_value * KDD]

    # Sum of KDDs only once per hot day
    total_kdd <- sum(hot_sample$KDD, na.rm = TRUE)

    # Average per offset using weighted mean
    avg_anomaly_by_offset <- boot_window_data[
      !is.na(weighted_anomaly),
      .(avg = sum(weighted_anomaly, na.rm = TRUE) / total_kdd),
      by = offset
    ]
    setkey(avg_anomaly_by_offset, offset)

    # Store in bootstrap matrix
    bootstrap_mat[i, match(avg_anomaly_by_offset$offset, day_offsets)] <- avg_anomaly_by_offset$avg
  }

  # --- Compute confidence intervals ---
  boot_summary <- data.table(
    day = day_offsets,
    mean = colMeans(bootstrap_mat, na.rm = TRUE),
    lower = apply(bootstrap_mat, 2, function(x) quantile(x, 0.025, na.rm = TRUE)),
    upper = apply(bootstrap_mat, 2, function(x) quantile(x, 0.975, na.rm = TRUE))
  )

  # --- Plot with ribbon ---
  p <- ggplot(boot_summary, aes(x = day)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), fill = "lightblue", alpha = 0.5) +
    geom_line(aes(y = mean), color = "blue", size = 1) +
    labs(
      title = paste("Bootstrapped 95% CI for", var, "(Weighted by Day 0 KDD)"),
      x = "Days from Hot Day (d0)",
      y = paste(var, "Anomaly")
    ) +
    theme_minimal()

  print(p)
  ggsave(
  filename = file.path("output_data", paste0(var, "_bootstrap_plot.png")),
  plot = p,
  width = 8,
  height = 5,
  dpi = 300
)
}


```


#4.0 Irrigated and Rainfed Yield Regressions

#4.1 Making a yearly dataframe out of PRISM climate data set
```{r}
# Summarize the dataset by GEOID and year


merged_prism_yearly <- merged_prism_full_1 %>%
  filter(month(time) %in% 5:9) %>%  # ✅ Filter for May (5) to September (9)
  select(-time, -binary_30, -KDD, -GDD) %>%  # Remove unwanted columns
  group_by(GEOID, year) %>%
  summarise(
    county_count = first(county_count),
    prism_tmax = mean(prism_tmax, na.rm = TRUE),
    prism_tmin = mean(prism_tmin, na.rm = TRUE),
    prism_vpdmean = mean(prism_vpdmean, na.rm = TRUE),
    prism_vpdmax = mean(prism_vpdmax, na.rm = TRUE),
    prism_vpdmin = mean(prism_vpdmin, na.rm = TRUE),
    prism_ppt = sum(prism_ppt, na.rm = TRUE),
    KDD_sum_yearly = first(KDD_sum_yearly),  # Already pre-summed
    GDD_sum_yearly = first(GDD_sum_yearly),
    # county = first(county_name),
    .groups = 'drop'
  )

```

#4.2 Load Yield Benefit Model data set
```{r}
# Read in datafrom environment
yield_benefit_model_data <- read_csv("input_data/yield-benefit-model-data.csv")
# Split the dataset by commodity_desc
crop_datasets <- split(yield_benefit_model_data, yield_benefit_model_data$commodity_desc)
list2env(crop_datasets, envir = .GlobalEnv)
```

#4.3
```{r}
# Merge CORN dataset with merged_prism_yearly by matching `fips` from CORN and `GEOID` from merged_prism_yearly
prism_CORN <- merge(CORN, merged_prism_yearly, by.x = c("fips", "year"), by.y = c("GEOID", "year"), all.x = TRUE)

#Merge SOYBEANS dataset with merged_prism_yearly by matching `fips` from SOYBEANS and `GEOID` from merged_prism_yearly
prism_SOYBEANS <- merge(SOYBEANS, merged_prism_yearly, by.x = c("fips", "year"), by.y = c("GEOID", "year"), all.x = TRUE)
```

#4.4 OPTIONAL: Filter for counties with 20 or more years of historical data
```{r}
# Filter prism_SOYBEANS
#prism_SOYBEANS <- prism_SOYBEANS %>%
 # group_by(fips) %>%
 # filter(n_distinct(year) >= 20) %>%
#  ungroup()

# Filter prism_CORN
#prism_CORN <- prism_CORN %>%
#  group_by(fips) %>%
#  filter(n_distinct(year) >= 20) %>%
#  ungroup()
```

#4.5 Convert county variable to a factor for later use
```{r}
# Convert county to a factor if it's not already a factor to treat it as a categorical variable
prism_SOYBEANS$county_factor <- as.factor(prism_SOYBEANS$fips) #repeat for corn

prism_CORN$county_factor <- as.factor(prism_CORN$fips) #repeat for corn

prism_CORN <- prism_CORN %>% 
  filter(util_practice_desc == "GRAIN")
```

#4.6 Run Linear Regressions for irrigated and rainfed yield for each crop
```{r}
# Linear regression for irrigated CORN yield
model_irrigated_CORN <- lm(irr~ KDD_sum_yearly + GDD_sum_yearly + prism_ppt, data = prism_CORN)
summary(model_irrigated_CORN)

# Linear regression for rainfed CORN yield
model_rainfed_CORN <- lm(rainfed ~ KDD_sum_yearly + GDD_sum_yearly + prism_ppt + county_factor, data = prism_CORN)
summary(model_rainfed_CORN)

# Linear regression for irrigated SOYBEANS yield
model_irrigated_SOYBEANS <- lm(irr~ KDD_sum_yearly + GDD_sum_yearly + prism_ppt, data = prism_SOYBEANS)
summary(model_irrigated_SOYBEANS)

# Linear regression for rainfed SOYBEANS yield
model_rainfed_SOYBEANS <- lm(rainfed ~ KDD_sum_yearly + GDD_sum_yearly + prism_ppt + county_factor, data = prism_SOYBEANS)
summary(model_rainfed_SOYBEANS)


#model_irrigated_interaction_ <- lm (irr ~ KDD_sum_yearly * year, data= prism_CORN)
#summary(model_irrigated_interaction)
```

#4.7 Running a Boostrap on the Crop Yield data to analyze the KDD coefficient 
```{r}

set.seed(123)

n_boot <- 1000

bootstrap_lm <- function(data, formula, n_boot = 1000) {
  base_model <- lm(formula, data = data)
  coef_names <- names(coef(base_model))

  boot_coefs <- replicate(n_boot, {
    boot_sample <- data[sample(nrow(data), replace = TRUE), ]
    result <- tryCatch({
      coefs <- coef(lm(formula, data = boot_sample))
      coefs_full <- setNames(rep(NA, length(coef_names)), coef_names)
      common <- intersect(names(coefs), coef_names)
      coefs_full[common] <- coefs[common]
      return(coefs_full)
    }, error = function(e) {
      setNames(rep(NA, length(coef_names)), coef_names)
    })
    return(result)
  }, simplify = TRUE)

  boot_df <- t(boot_coefs) %>% as.data.frame()
  return(boot_df)
}

crop_names <- c("corn", "soybeans")

boot_results <- list()

for (crop in crop_names) {
  crop_title <- toTitleCase(crop)
  data_name <- paste0("prism_", toupper(crop))
  
  if (!exists(data_name)) next
  crop_data <- get(data_name)
  
  for (water_type in c("rainfed", "irrigated")) {
    response <- ifelse(water_type == "rainfed", "rainfed", "irr")
    
    formula <- as.formula(paste0(response, " ~ KDD_sum_yearly + GDD_sum_yearly + prism_ppt + county_factor"))
    
    boot_df <- bootstrap_lm(crop_data, formula, n_boot)
    
    if (!"KDD_sum_yearly" %in% names(boot_df)) next
    
    kdd_coefs <- boot_df$KDD_sum_yearly
    kdd_coefs <- kdd_coefs[!is.na(kdd_coefs)]
    
    if (length(kdd_coefs) < 10) next
    
    boot_results[[length(boot_results) + 1]] <- data.frame(
      Crop = crop_title,
      Type = toupper(water_type),
      Mean = mean(kdd_coefs),
      Lower = quantile(kdd_coefs, 0.025),
      Upper = quantile(kdd_coefs, 0.975)
    )
  }
}

boot_results_df <- bind_rows(boot_results)

ggplot(boot_results_df, aes(x = Crop, y = Mean, color = Type)) +
  geom_point(size = 3, position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = Lower, ymax = Upper), width = 0.2, position = position_dodge(width = 0.5)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40") +
  scale_color_manual(values = c("IRRIGATED" = "blue", "RAINFED" = "green")) +
  labs(
    title = "Bootstrapped 95% CI for KDD Coefficient by Crop",
    x = "Crop",
    y = "KDD Coefficient (± 95% CI)",
    color = "Water Type"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```



# 4.8 Running a Boostrap on the Crop Yield data to analyze the KDD coefficient with 20+ years of historical data
```{r}
set.seed(123)
n_boot <- 1000

# --- Bootstrap helper function ---
bootstrap_lm <- function(data, formula, n_boot = 1000) {
  base_model <- lm(formula, data = data)
  coef_names <- names(coef(base_model))

  boot_coefs <- replicate(n_boot, {
    boot_sample <- data[sample(nrow(data), replace = TRUE), ]
    result <- tryCatch({
      coefs <- coef(lm(formula, data = boot_sample))
      coefs_full <- setNames(rep(NA, length(coef_names)), coef_names)
      common <- intersect(names(coefs), coef_names)
      coefs_full[common] <- coefs[common]
      return(coefs_full)
    }, error = function(e) {
      setNames(rep(NA, length(coef_names)), coef_names)
    })
    return(result)
  }, simplify = TRUE)

  boot_df <- t(boot_coefs) %>% as.data.frame()
  return(boot_df)
}

# --- Crop names and setup ---
crop_names <- c("corn", "soybeans")
boot_results <- list()

for (crop in crop_names) {
  crop_title <- toTitleCase(crop)
  data_name <- paste0("prism_", toupper(crop))
  
  if (!exists(data_name)) next
  crop_data <- get(data_name)
  
  # --- Filter to counties with ≥20 years of historical data ---
  crop_data <- crop_data %>%
    group_by(fips) %>%
    filter((max(year) - min(year) + 1) >= 20) %>%
    ungroup()
  
  for (water_type in c("rainfed", "irrigated")) {
    response <- ifelse(water_type == "rainfed", "rainfed", "irr")
    
    formula <- as.formula(
      paste0(response, " ~ KDD_sum_yearly + GDD_sum_yearly + prism_ppt + county_factor")
    )
    
    boot_df <- bootstrap_lm(crop_data, formula, n_boot)
    
    if (!"KDD_sum_yearly" %in% names(boot_df)) next
    
    kdd_coefs <- boot_df$KDD_sum_yearly
    kdd_coefs <- kdd_coefs[!is.na(kdd_coefs)]
    
    if (length(kdd_coefs) < 10) next
    
    boot_results[[length(boot_results) + 1]] <- data.frame(
      Crop = crop_title,
      Type = toupper(water_type),
      Mean = mean(kdd_coefs),
      Lower = quantile(kdd_coefs, 0.025),
      Upper = quantile(kdd_coefs, 0.975)
    )
  }
}

boot_results_df <- bind_rows(boot_results)

# --- Plot ---
ggplot(boot_results_df, aes(x = Crop, y = Mean, color = Type)) +
  geom_point(size = 3, position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = Lower, ymax = Upper),
                width = 0.2, position = position_dodge(width = 0.5)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40") +
  scale_color_manual(values = c("IRRIGATED" = "blue", "RAINFED" = "green")) +
  labs(
    title = "Bootstrapped 95% CI for KDD Coefficient by Crop (≥20 Years of Data per County)",
    x = "Crop",
    y = "KDD Coefficient (± 95% CI)",
    color = "Water Type"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

```{r}
library(dplyr)
library(ggplot2)

# --- Corn ---
corn_summary <- prism_CORN %>%
  group_by(fips) %>%
  summarise(n_years = n_distinct(year), .groups = "drop")

ggplot(corn_summary, aes(x = n_years)) +
  geom_histogram(binwidth = 1, fill = "blue", alpha = 0.7, color = "black") +
  labs(
    title = "Distribution of Number of Years per County - Corn",
    x = "Number of Years of Data",
    y = "Number of Counties"
  ) +
  theme_minimal()

# --- Soybeans ---
soy_summary <- prism_SOYBEANS %>%
  group_by(fips) %>%
  summarise(n_years = n_distinct(year), .groups = "drop")

ggplot(soy_summary, aes(x = n_years)) +
  geom_histogram(binwidth = 1, fill = "green", alpha = 0.7, color = "black") +
  labs(
    title = "Distribution of Number of Years per County - Soybeans",
    x = "Number of Years of Data",
    y = "Number of Counties"
  ) +
  theme_minimal()

```

