---
title: "Supplemental Figures: Climate Variables, Crop Distribution, and Historical Yields"
author: "Penny Hunter"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  warning = FALSE,
  message = FALSE
)
library(dplyr)
library(ggplot2)
library(sf)
library(tigris)
library(viridis)
library(patchwork)

options(tigris_use_cache = TRUE)
```

---

## Overview

This script builds Figure S1: county-level maps of mean growing-season
climate variables (GDD, KDD, precipitation, Tmax, VPDmax) and crop presence
across study counties (1981-2018), restricted to counties meeting the
minimum 10-year data requirement.

Panels a, b, d, and e (GDD, KDD, Tmax, VPDmax) use a yellow-to-purple
(magma) scale, where purple indicates higher values. Panel c (precipitation)
uses the standard viridis yellow-to-blue scale, where blue indicates wetter
conditions. Panel f shows crop presence by county.

---

## 1. Load Data

```{r load-data}
maize_gs    <- readRDS("input_data/maize_gs")
soybeans_gs <- readRDS("input_data/soybeans_gs")

counties_combined <- readRDS("input_data/counties_combined.rds")
# Remove Hidalgo County, NM (35031) - lacks USDA growing season data
counties_combined <- counties_combined %>%
  filter(fips != "35031")

counties_sf <- counties(cb = TRUE, class = "sf") %>%
  filter(!STATEFP %in% c("02", "15", "72", "78", "60", "66", "69"))

states_sf <- states(cb = TRUE, class = "sf") %>%
  filter(!STATEFP %in% c("02", "15", "72", "78", "60", "66", "69"))
```

---

## 2. Compute County-Level Climate Means

Mean growing-season GDD, KDD, precipitation, Tmax, and VPDmax are computed
per county, pooling across both crops and all years.

```{r compute-climate-means}
combine_climate <- function(df) {
  df %>%
    filter(fips %in% counties_combined$fips) %>%
    select(fips, KDD_sum_yearly, GDD_sum_yearly,
           prism_ppt, prism_tmax, prism_vpdmax)
}

climate_combined <- bind_rows(
  combine_climate(maize_gs),
  combine_climate(soybeans_gs)
) %>%
  distinct()

climate_summary <- climate_combined %>%
  group_by(fips) %>%
  summarise(
    mean_GDD    = mean(GDD_sum_yearly, na.rm = TRUE),
    mean_KDD    = mean(KDD_sum_yearly, na.rm = TRUE),
    mean_ppt    = mean(prism_ppt,      na.rm = TRUE),
    mean_tmax   = mean(prism_tmax,     na.rm = TRUE),
    mean_vpdmax = mean(prism_vpdmax,   na.rm = TRUE),
    .groups = "drop"
  )

map_data_climate <- counties_sf %>%
  mutate(fips = GEOID) %>%
  left_join(climate_summary, by = "fips")
```

---

## 3. Compute Crop Presence by County

```{r compute-crop-presence}
maize_counties <- maize_gs %>%
  filter(fips %in% counties_combined$fips) %>%
  pull(fips) %>%
  unique()

soy_counties <- soybeans_gs %>%
  filter(fips %in% counties_combined$fips) %>%
  pull(fips) %>%
  unique()

presence_df <- counties_combined %>%
  distinct(fips) %>%
  mutate(
    has_maize = fips %in% maize_counties,
    has_soy   = fips %in% soy_counties,
    Presence  = case_when(
      has_maize & has_soy   ~ "Both",
      has_maize & !has_soy  ~ "Maize only",
      !has_maize & has_soy  ~ "Soybean only",
      TRUE                  ~ NA_character_
    )
  ) %>%
  filter(!is.na(Presence))

map_data_presence <- counties_sf %>%
  mutate(fips = GEOID) %>%
  left_join(presence_df, by = "fips")
```

---

## 4. Plotting Function for Climate Panels

```{r plot-function}
plot_climate_panel <- function(data, states_sf, fill_var, panel_letter,
                                title, legend_name, palette = "magma",
                                direction = -1) {
  ggplot() +
    geom_sf(
      data = data,
      aes(fill = .data[[fill_var]]),
      color = NA
    ) +
    geom_sf(
      data = states_sf,
      fill = NA,
      color = "white",
      linewidth = 0.2
    ) +
    {
      if (palette == "magma") {
        scale_fill_viridis_c(
          option    = "magma",
          direction = direction,
          na.value  = "gray85",
          name      = legend_name
        )
      } else {
        scale_fill_viridis_c(
          option    = "viridis",
          direction = direction,
          na.value  = "gray85",
          name      = legend_name
        )
      }
    } +
    labs(title = paste0(panel_letter, ". ", title)) +
    coord_sf(datum = NA) +
    theme_minimal(base_size = 8, base_family = "Helvetica") +
    theme(
      axis.text       = element_blank(),
      axis.ticks      = element_blank(),
      panel.grid      = element_blank(),
      plot.title      = element_text(face = "bold", size = 9, hjust = 0),
      legend.title    = element_text(size = 7),
      legend.text     = element_text(size = 6),
      legend.position = "bottom",
      legend.key.height = unit(0.3, "cm"),
      legend.key.width  = unit(0.6, "cm")
    )
}
```

---

## 5. Build Panels a-e (Climate Variables)

GDD, KDD, Tmax, and VPDmax use the magma palette (yellow to purple, purple =
higher). Precipitation uses the standard viridis palette (yellow to blue,
blue = wetter).

```{r build-climate-panels}
p_gdd <- plot_climate_panel(
  map_data_climate, states_sf, "mean_GDD", "a",
  "Mean Growing Degree Days", "\u00b0C\u00b7day",
  palette = "magma", direction = -1
)

p_kdd <- plot_climate_panel(
  map_data_climate, states_sf, "mean_KDD", "b",
  "Mean Killing Degree Days", "\u00b0C\u00b7day",
  palette = "magma", direction = -1
)

p_ppt <- plot_climate_panel(
  map_data_climate, states_sf, "mean_ppt", "c",
  "Mean Precipitation", "mm",
  palette = "viridis", direction = -1
)

p_tmax <- plot_climate_panel(
  map_data_climate, states_sf, "mean_tmax", "d",
  "Mean Tmax", "\u00b0C",
  palette = "magma", direction = -1
)

p_vpd <- plot_climate_panel(
  map_data_climate, states_sf, "mean_vpdmax", "e",
  "Mean VPD Max", "kPa",
  palette = "magma", direction = -1
)
```

---

## 6. Build Panel f (Crop Presence)

```{r build-presence-panel}
p_presence <- ggplot() +
  geom_sf(
    data = map_data_presence,
    aes(fill = Presence),
    color = NA
  ) +
  geom_sf(
    data = states_sf,
    fill = NA,
    color = "white",
    linewidth = 0.2
  ) +
  scale_fill_manual(
    values = c(
      "Both"         = "#1f78b4",
      "Maize only"   = "#f0c419",
      "Soybean only" = "#33a02c"
    ),
    na.value = "gray85",
    name     = NULL
  ) +
  labs(title = "f. Dataset Representation by Crop and County") +
  coord_sf(datum = NA) +
  theme_minimal(base_size = 8, base_family = "Helvetica") +
  theme(
    axis.text       = element_blank(),
    axis.ticks      = element_blank(),
    panel.grid      = element_blank(),
    plot.title      = element_text(face = "bold", size = 9, hjust = 0),
    legend.text     = element_text(size = 7),
    legend.position = "bottom"
  )
```

---

## 7. Combine Panels into Figure S1

```{r combine-figure-s1, fig.width=7.5, fig.height=9}
figure_s1 <- (p_gdd | p_kdd) /
  (p_ppt  | p_tmax) /
  (p_vpd  | p_presence)

figure_s1

ggsave(
  filename = "output_data/Figure_S1.png",
  plot     = figure_s1,
  width    = 7.5,
  height   = 9,
  dpi      = 600,
  units    = "in",
  bg       = "white"
)
```
